---
title: Niche overlap increases with increasing resource availability in a tropical
  plant-pollinator network.
author: "Ben Weinstein"
date: "Tuesday, October 14, 2014"
output:
  html_document:
    theme: spacelab
  word_document: default
---
```{r,echo=F}
paste("Run Completed at",Sys.time())
```

```{r,echo=FALSE,warning=FALSE,message=FALSE}
require(knitr)
library(R2jags)
require(reshape2)
require(ggplot2)
library(scales)
library(dplyr)
library(stringr)
require(picante)
require(igraph)
require(bipartite)
library(chron)
library(gridExtra)
library(doSNOW)
library(foreach)
#Set Dropbox Location
#setwd to dropbox
droppath<-"C:/Users/Ben/Dropbox/"

#Set github path
gitpath<-"C:/Users/Ben/Documents/NetworkTime/"

setwd(gitpath)

opts_chunk$set(message=FALSE,warning=FALSE,fig.height=7,fig.width=11,echo=FALSE,cache=TRUE,cache.path = 'NetworkTime_cache/', fig.path='figure/')

```

************

```{r,fig.height=5,fig.width=8}
#Get flower transect data
full.fl<-read.csv("C:/Users/Ben/Dropbox/Thesis/Maquipucuna_SantaLucia/Results/FlowerTransects/FlowerTransectClean.csv")[,-1]

 #month should be capital 
colnames(full.fl)[colnames(full.fl) %in% "month"]<-"Month"

#group by month and replicate, remove date errors by making a max of 10 flowers, couple times where the gps places it in wrong transect by 1 to 2 meters. 
flower.month<-group_by(full.fl,Month,Year,Transect_R,Date_F) %>% dplyr::summarise(Flowers=sum(Total_Flowers,na.rm=TRUE))  %>% filter(Flowers>20)
  
#Make month abbreviation column, with the right order
flower.month$Month.a<-factor(month.abb[flower.month$Month],month.abb[c(1:12)])

#Make year factor column
flower.month$Year<-as.factor(flower.month$Year)

#get quantile for each transect
thresh<-melt(group_by(flower.month,Transect_R) %>% summarize(Threshold=quantile(log(Flowers),0.5)) )
flower.month<-merge(flower.month,thresh)
flower.month$High<-log(flower.month$Flowers)>flower.month$value

#fix the levels
levels(flower.month$Transect_R)<-c("1300m - 1500m", "1500m - 1700m","1700m - 1900m","1900m - 2100m","2100m - 2300m","2300m - 2500m")
#plot
ggplot(flower.month,aes(x=Month.a,log(Flowers),col=High,shape=as.factor(Year))) + geom_point(size=3) + theme_bw()  + geom_smooth(aes(group=1)) + ylab("Flowers") + xlab("Month") + facet_wrap(~Transect_R,scales="free_y") + labs(shape="Year", y= "Log Available Flowers") + scale_x_discrete(breaks=month.abb[seq(1,12,2)]) + scale_color_manual(labels=c("Low","High"),values=c("black","red")) + labs(col="Resource Availability")

ggsave("Figures/AvailableFlowers.jpg",dpi=600,units="in",width=7,height=6)
ggsave("Figures/AvailableFlowers.svg",units="in",width=7,height=6)

```

****

**Figure 3. Measures of hummingbird plant interactions through time (points) as compared to a null model maintaining the number of observations (shaded region).The size of the point is proportional to the number of interactions measured for each month.**

#Read in data

From Generate.Rmd

```{r,fig.height=10,fig.width=11}

int<-read.csv("InputData/FinalObs.csv",row.names=1)

#just presence records
int<-int[int$State=="Used",]

int$Year<-years(int$DateP)
int$Year[int$Year %in% 2012]<-2013

int[int$Year %in% 2013 & int$Month %in% 1,"Year"]<-2014

#break up network by two months
int.M<-split(int,list(int$Month,int$Year),drop=TRUE)

#remove months with less than 15 records
int.M<-int.M[sapply(int.M,nrow)>15]

#design null function
nullC<-function(x){
  x<-droplevels(x)
  N=nrow(x)
  month.inter<-table(x$Hummingbird,x$Iplant_Double)

true_stat<-networklevel(month.inter,level="lower",index=c("weighted connectance","weighted NODF"))
  
true_stat<-melt(true_stat)

true_stat$variable<-rownames(true_stat)

# Create 199 random matrixes
r.int<-permatfull(month.inter,times=500,fixed="both")

#Create a function that computes network metrics
null_total<-as.data.frame(
  t(sapply(r.int$perm,function(x){
  #Compute network statistics
    out<-networklevel(x,level="lower",index=c("weighted connectance","weighted NODF"))
    return(out)
    })))
#get the upper and lower 2.5th quantiles
null_total<-melt(null_total)
nullq<-group_by(null_total,variable) %>% summarize(Lower=quantile(value,0.025,na.rm=T),Upper=quantile(value,0.975,na.rm=T))

#add month
nullq$Month<-unique(x$Month)

#add Year
nullq$Year<-unique(x$Year)

colnames(nullq)<-c("Metric","Lower","Upper","Month","Year")

#append true information
out<-merge(nullq,true_stat,by.x="Metric",by.y="variable")
colnames(out)[colnames(out) %in% "value"]<-"True"
out$N<-N
return(out)}

#apply function to each month
cl<-makeCluster(2,"SOCK")
registerDoSNOW(cl)
runs<-foreach(x=1:length(int.M),.packages=c("bipartite","reshape2","dplyr")) %dopar% {nullC(int.M[[x]])}
stopCluster(cl)

datHL<-rbind_all(runs)

#format date
datHL$Date<-as.Date(paste(1,round(datHL$Month),datHL$Year,sep="/"),format="%d/%m/%Y")

flowerMean<-group_by(flower.month,Month,Year,Transect_R) %>% summarize(mFlowers=mean(Flowers,na.rm=TRUE)) %>% group_by(Month,Year) %>% summarize(Flowers=sum(mFlowers,na.rm=TRUE))
  
#combine the flower totals and network metrics
  network.fl<-merge(flowerMean,datHL,by=c("Month","Year"))

#plot distribution of  values
p<-ggplot(network.fl,aes(x=Date)) + facet_wrap(~Metric,scales='free',ncol=1) + geom_ribbon(aes(ymin=Lower,ymax=Upper),alpha=0.3)
p<-p+ geom_point(aes(size=N,y=True,col=log(Flowers))) + geom_line(aes(y=True)) 
p<-p+scale_color_continuous(low='blue',high='red') + scale_size_continuous(range=c(2.5,5),"Observations") + labs(col="Log Flower Abundance",y="") + theme_bw() + scale_x_date(date_breaks="3 months",date_labels = "%m/%Y") 
print(p) 

ggsave("Figures/NullNetwork.jpg",dpi=600,quality=100,pointsize=14,height=6,width=8)
ggsave("Figures/NullNetwork.svg",height=10,width=12)

```

********

**Figure 4. Connectance, clustering and niche overlap as a function of available resources. Point size is proportional to number of interactions observed that month.**

```{r,fig.height=10,fig.width=12}
    
  #Quick visualization 
  p<-ggplot(network.fl[,],aes(log(Flowers),y=True,shape=as.factor(Year))) + geom_point(aes(size=N)) + facet_wrap(~Metric,scale="free",ncol=1)  + geom_smooth(aes(group=1),method="lm") + theme_bw() + labs(col="Month",shape="Year",size="Interactions") + geom_text(aes(label=month.abb[Month]),hjust=-0.25,size=3) + scale_size_continuous(range=c(2.5,5),"Observations") + labs(x="Log Available Resources") + ylab("Score")
p
ggsave("Figures/Network_Resources.jpg",height=6,width=6,dpi=300,units="in")
ggsave("Figures/Network_Resources.svg",height=10,width=10)

```

Regression
```{r}
network.fl$value<-network.fl$True
summary(glm(data=network.fl[network.fl$Metric %in% "weighted NODF",],value~log(Flowers)),family="binomial")
```

