---
title: Appendix  B - Network Statistics.
author: "Ben Weinstein"
date: "2/1/2016"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
    keep_md: true
---

```{r,echo=F}
paste("Run Completed at",Sys.time())
```

```{r,echo=FALSE,warning=FALSE,message=FALSE}
require(knitr)
library(R2jags)
require(reshape2)
require(ggplot2)
library(scales)
library(dplyr)
library(stringr)
require(picante)
require(igraph)
require(bipartite)
library(chron)
library(gridExtra)
library(doSNOW)
library(foreach)
#Set Dropbox Location
#setwd to dropbox
droppath<-"C:/Users/Ben/Dropbox/"

#Set github path
gitpath<-"C:/Users/Ben/Documents/NetworkTime/"

setwd(gitpath)

opts_chunk$set(message=FALSE,warning=FALSE,fig.height=5,fig.width=10,echo=FALSE,cache=F,cache.path = 'NetworkTime_cache/', fig.path='figure/',fig.align = 'center')

```


************

```{r}
#load data if needed
#load("NetworkTime.Rdata")
```

#Read in data

From Generate.Rmd

```{r,fig.height=20,fig.width=13}

int<-read.csv("InputData/FinalObs.csv",row.names=1)

#just presence records
#int<-int[int$State=="Used",]

#Just use cameras
int<-int[int$Survey_Type=="Camera",]

fl.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/FlowerMorphology.csv",sep=""))

#First row is empty
fl.morph<-fl.morph[-1,]

#Bring in Hummingbird Morphology Dataset, comes from
hum.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/HummingbirdMorphology.csv",sep=""))

#taxonomy change, we are calling them Crowned Woodnymph's now.
hum.morph$English<-as.character(hum.morph$English)
hum.morph$English[hum.morph$English %in% "Green-crowned Woodnymph"]<-"Crowned Woodnymph"
```

#Daily interaction rate
```{r}
ndays<-int %>% filter(State=="Used") %>% group_by(Iplant_Double) %>% summarize(n=length(unique(DateP)))

#Total visits 
tvisits<-int %>% filter(State=="Used") %>% group_by(Hummingbird,Iplant_Double) %>% summarize(Total_Visits=sum(Yobs))
tvisits<-merge(tvisits,ndays,by="Iplant_Double")
tvisits$Rate<-round(tvisits$Total_Visits/tvisits$n,3)

```

#Overall Network

## Ordered by interactions

```{r}
#get order of interactions
hord<-tvisits %>% group_by(Hummingbird) %>% summarize(n=sum(Rate)) %>% arrange(n)

tvisits$Hummingbird<-factor(tvisits$Hummingbird,levels=hord$Hummingbird)

pord<-tvisits %>% group_by(Iplant_Double) %>% summarize(n=sum(Rate)) %>% arrange(desc(n))

tvisits$Iplant_Double<-factor(tvisits$Iplant_Double,levels=pord$Iplant_Double)

overall<-ggplot(tvisits,aes(y=Hummingbird,x=Iplant_Double,fill=Rate)) + theme_bw() + geom_tile() + scale_fill_continuous(low="blue",high="red",na.value = "white") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Plant") 
overall

#cast web
tweb<-acast(tvisits,Hummingbird~Iplant_Double,value.var="Rate",fill=0)
```

##Ordered by traits

```{r}
#get order of interactions by bill 
hord<-hum.morph[,c("English","Total_Culmen")] %>% arrange(desc(Total_Culmen)) %>% .$English

#by corolla
pord<-fl.morph %>% arrange(TotalCorolla) %>% .$Group.1

tvisits$Hummingbird<-factor(tvisits$Hummingbird,levels=hord)
tvisits$Iplant_Double<-factor(tvisits$Iplant_Double,levels=pord)


ggplot(tvisits,aes(y=Hummingbird,x=Iplant_Double,fill=Rate)) + theme_bw() + geom_tile() + scale_fill_continuous(low="blue",high="red",na.value = "white") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Plant") 

```

##Calculate overall statistics.

```{r}
overallmetrics<-networklevel(as.matrix(tweb),level="lower",index=c("weighted connectance","weighted NODF"))
#melt into data frame
overallmetrics<-data.frame(Metric=names(overallmetrics),value=overallmetrics)
```

## Permutation tests

We want to randomize the underlying total matrix, then recalculate daily interaction rates. We maintain column sums since cameras were placed at individual flowers, those identities are fixed through sampling. It is only the identity of hummingbirds that could vary in create realistic alternative networks. 

### Calculate randomized matrix

Maintaining the total number of observations, but shuffling where those visitations fall.

```{r}

null_metrics<-list()
for(x in 1:500){
  
  #create duplicate dataframe
  nint<-int
  
  #Resample hummingbird identity based on elevation
  nint$State<-sample(nint$State)
  
  #Calculate Daily interaction rate
  ndays<-nint %>% filter(State=="Used") %>% group_by(Iplant_Double) %>% summarize(n=length(unique(DateP))) 

  #Total visits 
  nvisits<-nint %>% filter(State=="Used") %>% group_by(Hummingbird,Iplant_Double) %>% summarize(Total_Visits=sum(Yobs))
  nvisits<-merge(nvisits,ndays,by="Iplant_Double")
  nvisits$Rate<-round(nvisits$Total_Visits/nvisits$n,3)
  
  #cast into network
  nweb<-acast(nvisits,Hummingbird~Iplant_Double,value.var="Rate",fill=0)
  nm<-networklevel(nweb,level="lower",index=c("weighted connectance","weighted NODF"))
  null_metrics[[x]]<-data.frame(Metric=names(nm),value=nm)
}

null_metrics<-melt(null_metrics)
ggplot(data=null_metrics,aes(x=value)) + geom_density(fill='grey40') + facet_wrap(~Metric,scales="free") + geom_vline(data=overallmetrics,aes(xintercept=value),col='red',linetype='dashed') + theme_bw()
```

```{r}
#Add overall network and measures underneath as a supp figure
num<-droplevels(null_metrics[null_metrics$Metric %in% c( "weighted connectance","weighted NODF"),])

ob<-droplevels(overallmetrics[overallmetrics$Metric %in% c("weighted connectance","weighted NODF"),])

levels(num$Metric)<-c("Connectance","Nestedness")
levels(ob$Metric)<-c("Connectance","Nestedness")

nulls<-ggplot(num,aes(x=value)) + geom_density(fill='grey40',alpha=0.5) + facet_wrap(~Metric,scales="free",nrow=1) + geom_vline(data=ob,aes(xintercept=value),col='red',linetype='dashed',size=1.5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

jpeg("Figures/Overall.jpeg",height=6,width=11,units="in",res=400)
grid.arrange(overall,nulls,heights=c(.7,.3)) 
dev.off()
```

#Temporal change in network

##View Sub-Networks

```{r,fig.height=15}

int.M<-split(int,list(int$Month,int$Year),drop = T)

p<-lapply(int.M,function(x){

  days<-x %>% filter(State=="Used") %>% group_by(Iplant_Double) %>% summarize(n=length(unique(DateP))) 

  #Total visits 
  tvisits<-x %>% filter(State=="Used") %>% group_by(Hummingbird,Iplant_Double) %>% summarize(Total_Visits=sum(Yobs))
  tvisits<-merge(tvisits,ndays,by="Iplant_Double")
  tvisits$Rate<-round(tvisits$Total_Visits/tvisits$n,3)

  tweb<-acast(tvisits,Hummingbird~Iplant_Double,value.var="Rate",fill=0)
  
  #turn 0's to NA for plotting
  tvisits$Rate[tvisits$Rate==0]<-NA
  tvisits<-tvisits[!is.na(tvisits$Rate),]
  
  #get order of interactions
  hord<-tvisits %>% group_by(Hummingbird) %>% filter(!is.na(Rate)) %>% summarize(n=sum(Rate)) %>% arrange(n) %>% .$Hummingbird
  
  tvisits$Hummingbird<-factor(tvisits$Hummingbird,levels=hord)
  
  pord<-tvisits %>% group_by(Iplant_Double) %>% filter(!is.na(Rate)) %>% summarize(n=sum(Rate)) %>% arrange(desc(n)) %>% .$Iplant_Double
  
  tvisits$Iplant_Double<-factor(tvisits$Iplant_Double,levels=pord)
  
  ggplot(tvisits,aes(y=Hummingbird,x=Iplant_Double,fill=Rate)) + theme_bw() +geom_tile() + scale_fill_continuous(low="blue",high="red",na.value = "white") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Plant") + ggtitle(unique(x$Availability)) 
})

p
```


## Randomize with respect to sampling

```{r}

make_web<-function(x){
    #create duplicate dataframe
  nint<-x

  #Calculate Daily interaction rate
  ndays<-nint %>% filter(State=="Used") %>% group_by(Iplant_Double) %>% summarize(n=length(unique(DateP))) 

  #Total visits 
  nvisits<-nint %>% filter(State=="Used") %>% group_by(Hummingbird,Iplant_Double) %>% summarize(Total_Visits=sum(Yobs))
  nvisits<-merge(nvisits,ndays,by="Iplant_Double")
  nvisits$Rate<-round(nvisits$Total_Visits/nvisits$n,3)
  
  #cast into network
  nweb<-acast(nvisits,Hummingbird~Iplant_Double,value.var="Rate",fill=0)
  return(nweb)
}

null_month<-function(x){
  
  #make duplicate
  nint<-x
  
  #make web
  tweb<-make_web(nint)
  true_state<-melt(networklevel(tweb,level="lower",index=c("weighted connectance","weighted NODF")))
  true_state<-data.frame(Metric=rownames(true_state),true_value=true_state,Availability=unique(nint$Availability))
  
null_state<-list()
  for(i in 1:100){
    #make duplicate, poor scoping.
    nint<-x
    nint$State<-sample(nint$State)
    nweb<-make_web(nint)
    ns<-melt(networklevel(nweb,level="lower",index=c("weighted connectance","weighted NODF")))
    null_state[[i]]<-data.frame(Metric=rownames(ns),value=ns)
  }
  
  #get the upper and lower 5th quantiles
  null_total<-melt(null_state,id.vars="Metric")
  nullq<-group_by(null_total,Metric) %>% summarize(Lower=quantile(value,0.05,na.rm=T),Upper=quantile(value,0.95,na.rm=T))
  
  #append true information
  out<-merge(nullq,true_state,by.x="Metric")
  colnames(out)[colnames(out) %in% "value"]<-"True_State"
return(out)}
```

##Which months were sufficiently sampled?

```{r}
#need to do some rarifaction?
#atleast 20 observations for now.

samp<-sapply(int.M,function(x){
nrow(x[x$State=="Used",])
  })
int.M<-int.M[samp>15]
```

###Apply function to each period

```{r}
#needs to be fixed
runs<-list()
for(x in 1:length(int.M)){
  runs[[x]]<-null_month(int.M[[x]])
}

names(runs)<-names(int.M)
datHL<-melt(runs,id.vars=colnames(runs[[1]]))

#Create time stamp
datHL[,c("Month","Year")]<-str_match(datHL$L1,"(\\d+).(\\d+)")[,-1]
datHL$TimeStamp<-strptime(paste("01",datHL$Month,datHL$Year,sep="/"),format="%d/%m/%Y")
```

#Observed versus null values of nestedness, connectance within time periods

```{r}
p<-ggplot(datHL,aes(x=TimeStamp,fill=Metric)) + facet_wrap(~Metric,scales='free',ncol=1) + geom_linerange(aes(ymin=Lower,ymax=Upper),col="red")
p<-p+ geom_point(aes(y=True_State,col=Availability,size=3)) + geom_line(aes(y=True_State)) 
p<-p+scale_fill_discrete(guide="none") + labs(y="") + theme_bw()
print(p) 
ggsave("Figures/NullNetwork.jpg",dpi=600,quality=100,pointsize=14,height=10,width=12)
ggsave("Figures/NullNetwork.svg",height=10,width=12)

```

```{r}
save.image("NetworkTime.RData")
```
