---
title: Appendix  B - Species Network Statistics.
author: "Ben Weinstein"
date: "2/1/2016"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
    keep_md: true
---

```{r,echo=F}
paste("Run Completed at",Sys.time())
```

```{r,echo=FALSE,warning=FALSE,message=FALSE}
require(knitr)
library(R2jags)
require(reshape2)
require(ggplot2)
library(scales)
library(dplyr)
library(stringr)
require(picante)
require(igraph)
require(bipartite)
library(chron)
library(gridExtra)
library(doSNOW)
library(foreach)
#Set Dropbox Location
#setwd to dropbox
droppath<-"C:/Users/Ben/Dropbox/"

#Set github path
gitpath<-"C:/Users/Ben/Documents/NetworkTime/"

setwd(gitpath)

opts_chunk$set(message=FALSE,warning=FALSE,fig.height=5,fig.width=12,echo=FALSE,cache=F,cache.path = 'NetworkTime_cache/', fig.path='figure/',fig.align = 'center')

```


************

```{r}
#load data if needed
#load("NetworkTime.Rdata")
```

#Read in data


```{r,fig.height=20,fig.width=13}
#From Generate.Rmd

int<-read.csv("InputData/FinalObs.csv",row.names=1)

#just presence records
#int<-int[int$State=="Used",]

#Just use cameras
int<-int[int$Survey_Type=="Camera",]

fl.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/FlowerMorphology.csv",sep=""))

#First row is empty
fl.morph<-fl.morph[-1,]

#Bring in Hummingbird Morphology Dataset, comes from
hum.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/HummingbirdMorphology.csv",sep=""))

#taxonomy change, we are calling them Crowned Woodnymph's now.
hum.morph$English<-as.character(hum.morph$English)
hum.morph$English[hum.morph$English %in% "Green-crowned Woodnymph"]<-"Crowned Woodnymph"
```

#Daily interaction rate
```{r}
ndays<-int %>% filter(State=="Used") %>% group_by(Iplant_Double) %>% summarize(n=length(unique(DateP)))

#Total visits 
tvisits<-int %>% filter(State=="Used") %>% group_by(Hummingbird,Iplant_Double) %>% summarize(Total_Visits=sum(Yobs))
tvisits<-merge(tvisits,ndays,by="Iplant_Double")
tvisits$Rate<-round(tvisits$Total_Visits/tvisits$n,3)
```

#Overall Network

## Ordered by interactions

```{r}
#get order of interactions
hord<-tvisits %>% group_by(Hummingbird) %>% summarize(n=sum(Rate)) %>% arrange(n)

tvisits$Hummingbird<-factor(tvisits$Hummingbird,levels=hord$Hummingbird)

pord<-tvisits %>% group_by(Iplant_Double) %>% summarize(n=sum(Rate)) %>% arrange(desc(n))

tvisits$Iplant_Double<-factor(tvisits$Iplant_Double,levels=pord$Iplant_Double)

overall<-ggplot(tvisits,aes(y=Hummingbird,x=Iplant_Double,fill=Rate)) + theme_bw() + geom_tile() + scale_fill_continuous(low="blue",high="red",na.value = "white") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Plant") 
overall

#cast web
tweb<-acast(tvisits,Hummingbird~Iplant_Double,value.var="Rate",fill=0)
```

##Ordered by traits

```{r}
#get order of interactions by bill 
hord<-hum.morph[,c("English","Total_Culmen")] %>% arrange(desc(Total_Culmen)) %>% .$English

#by corolla
pord<-fl.morph %>% arrange(TotalCorolla) %>% .$Group.1

tvisits$Hummingbird<-factor(tvisits$Hummingbird,levels=hord)
tvisits$Iplant_Double<-factor(tvisits$Iplant_Double,levels=pord)

ggplot(tvisits,aes(y=Hummingbird,x=Iplant_Double,fill=Rate)) + theme_bw() + geom_tile() + scale_fill_continuous(low="blue",high="red",na.value = "white") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Plant") 
```

##Calculate overall statistics.

```{r}
overallmetrics<-specieslevel(as.matrix(tweb),level="lower",index=c("PDI"))
#melt into data frame
overallmetrics<-data.frame(Hummingbird=rownames(overallmetrics),Metric=names(overallmetrics),value=overallmetrics)
```

## Permutation tests

We want to randomize the underlying total matrix, then recalculate daily interaction rates. We maintain column sums since cameras were placed at individual flowers, those identities are fixed through sampling. It is only the identity of hummingbirds that could vary in create realistic alternative networks. 

### Calculate randomized matrix

Maintaining the total number of observations, but shuffling where those visitations fall.

```{r}
null_metrics<-list()
for(x in 1:100){
  
  #create duplicate dataframe
  nint<-int
  
  #Resample hummingbird identity based on elevation
  nint$Hummingbird<-sample(nint$Hummingbird)
  
  #Calculate Daily interaction rate
  ndays<-nint %>% filter(State=="Used") %>% group_by(Iplant_Double) %>% summarize(n=length(unique(DateP))) 

  #Total visits 
  nvisits<-nint %>% filter(State=="Used") %>% group_by(Hummingbird,Iplant_Double) %>% summarize(Total_Visits=sum(Yobs))
  nvisits<-merge(nvisits,ndays,by="Iplant_Double")
  nvisits$Rate<-round(nvisits$Total_Visits/nvisits$n,3)
  
  #cast into network
  nweb<-acast(nvisits,Hummingbird~Iplant_Double,value.var="Rate",fill=0)
  nm<-specieslevel(nweb,level="lower",index=c("PDI"))
  null_metrics[[x]]<-data.frame(Hummingbird=rownames(overallmetrics),Metric=names(nm),value=nm)
}

hord<-hum.morph[,c("English","Total_Culmen")] %>% arrange(desc(Total_Culmen)) %>% .$English
null_metrics$Hummingbird<-factor(null_metrics$Hummingbird,levels=hord)

null_metrics<-melt(null_metrics)
ggplot(data=null_metrics,aes(x=Hummingbird,y=value)) + geom_violin(fill='grey40',alpha=0.5) + geom_point(data=overallmetrics,aes(x=Hummingbird,y=PDI),col='red',size=2.5) + theme_bw() + coord_flip()
```

```{r}
nulls<-ggplot(null_metrics,aes(x=value)) + geom_density(fill='grey40',alpha=0.5) + facet_wrap(~Metric,scales="free",nrow=1) + geom_vline(data=overallmetrics,aes(xintercept=PDI),col='red',linetype='dashed',size=1.5) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

jpeg("Figures/Overall.jpeg",height=6,width=11,units="in",res=400)
grid.arrange(overall,nulls,heights=c(.7,.3)) 
dev.off()
```

#Temporal change in network

##View Sub-Networks

```{r,fig.height=40,fig.width=12}
#Create time stamp
int$TimeStamp<-as.POSIXct(int$DateP)

#Set window size
cutm<-"2 month"
int$Date_C<-cut(int$TimeStamp,breaks=cutm)
int.M<-split(int,int$Date_C,drop = T)

webp<-lapply(int.M,function(x){

  days<-x %>% filter(State=="Used") %>% group_by(Iplant_Double) %>% summarize(n=length(unique(DateP))) 
  #Total visits 
  tvisits<-x %>% filter(State=="Used") %>% group_by(Hummingbird,Iplant_Double) %>% summarize(Total_Visits=sum(Yobs))
  tvisits<-merge(tvisits,ndays,by="Iplant_Double")
  tvisits$Rate<-round(tvisits$Total_Visits/tvisits$n,3)

  tweb<-acast(tvisits,Hummingbird~Iplant_Double,value.var="Rate",fill=0)
  
  #turn 0's to NA for plotting
  tvisits$Rate[tvisits$Rate==0]<-NA
  tvisits<-tvisits[!is.na(tvisits$Rate),]
  
  #get order of interactions
  hord<-tvisits %>% group_by(Hummingbird) %>% filter(!is.na(Rate)) %>% summarize(n=sum(Rate)) %>% arrange(n) %>% .$Hummingbird
  
  tvisits$Hummingbird<-factor(tvisits$Hummingbird,levels=hord)
  
  pord<-tvisits %>% group_by(Iplant_Double) %>% filter(!is.na(Rate)) %>% summarize(n=sum(Rate)) %>% arrange(desc(n)) %>% .$Iplant_Double
  
  tvisits$Iplant_Double<-factor(tvisits$Iplant_Double,levels=pord)
  
  ggplot(tvisits,aes(y=Hummingbird,x=Iplant_Double,fill=Rate)) + theme_bw() +geom_tile() + scale_fill_continuous(low="blue",high="red",na.value = "white") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Plant") + ggtitle(unique(x$Date_C)) 
})
marrangeGrob(webp,nrow=8,ncol=2)
```

## Randomize with respect to sampling

```{r}
make_web<-function(x){
    #create duplicate dataframe
  nint<-x

  #Calculate Daily interaction rate
  ndays<-nint %>% filter(State=="Used") %>% group_by(Iplant_Double) %>% summarize(n=length(unique(DateP))) 

  #Total visits 
  nvisits<-nint %>% filter(State=="Used") %>% group_by(Hummingbird,Iplant_Double) %>% summarize(Total_Visits=sum(Yobs))
  nvisits<-merge(nvisits,ndays,by="Iplant_Double")
  nvisits$Rate<-round(nvisits$Total_Visits/nvisits$n,3)
  
  #cast into network
  nweb<-acast(nvisits,Hummingbird~Iplant_Double,value.var="Rate",fill=0)
  return(nweb)
}

null_month<-function(x){
  
  #make duplicate
  nint<-x
  
  #make web
  tweb<-make_web(nint)
  true_state<-specieslevel(tweb,level="lower",index="PDI")
  true_state<-data.frame(Metric=rownames(true_state),true_value=true_state)
  
null_state<-list()
  for(i in 1:100){
    #make duplicate
    nint<-x
    #resample hummingbird identity
    nint$Hummingbird<-sample(nint$Hummingbird)
    nweb<-make_web(nint)
    ns<-specieslevel(nweb,level="lower",index=c("PDI"))
    null_state[[i]]<-data.frame(Metric=rownames(ns),value=ns)
  }
  
  #get the upper and lower 5th quantiles
  null_total<-melt(null_state,id.vars="Metric")
  nullq<-group_by(null_total,Metric) %>% summarize(Mean=mean(value),Lower=quantile(value,0.05,na.rm=T),Upper=quantile(value,0.95,na.rm=T))
  
  #append true information
  out<-merge(nullq,true_state,by.x="Metric")
  colnames(out)[colnames(out) %in% "value"]<-"True_State"
return(out)}
```


```{r}
##Which months were sufficiently sampled?
#need to do some rarifaction?
#atleast 20 observations for now.

samp<-sapply(int.M,function(x){
nrow(x[x$State=="Used",])
  })

```

###Apply function to each period

```{r}
  #needs to be fixed
runs<-list()
for(x in 1:length(int.M)){
  runs[[x]]<-null_month(int.M[[x]])
}

names(runs)<-names(int.M)
datHL<-melt(runs,id.vars=colnames(runs[[1]]))

#Create time stamp
datHL[,c("Year","Month")]<-str_match(datHL$L1,"(\\d+).(\\d+)")[,-1]
datHL$Date_C<-strptime(paste("01",datHL$Month,datHL$Year,sep="/"),format="%d/%m/%Y")
```

```{r}
#Merge Flower Data

#Get flower transect data
full.fl<-read.csv("C:/Users/Ben/Dropbox/Thesis/Maquipucuna_SantaLucia/Results/FlowerTransects/FlowerTransectClean.csv")[,-1]

#month should be capital 
colnames(full.fl)[colnames(full.fl) %in% "month"]<-"Month"

#Take out flowers not in the network
full.fl<-full.fl[full.fl$Iplant_Double %in% levels(int$Iplant_Double),]

#Time Stamp
full.fl$Date_F<-as.POSIXct(full.fl$Date_F)

#Cut Stamp
full.fl$Date_C<-cut(full.fl$Date_F,breaks=unique(datHL$Date_C))

#group by date cut
flower.month<-group_by(full.fl,Date_C,Transect_R,Date) %>% dplyr::summarise(Flowers=sum(Total_Flowers,na.rm=TRUE))%>% filter(Flowers>10) %>% summarize(Flowers=mean(Flowers)) %>% summarize(Flowers=sum(Flowers))

#combine the flower totals and network metrics
network.fl<-merge(flower.month,datHL,by="Date_C")
```

```{r}
p<-ggplot(network.fl,aes(x=as.POSIXct(Date_C),fill=Metric)) + facet_wrap(~Metric,scales='free',ncol=4) + geom_linerange(aes(ymin=Lower,ymax=Upper),col="red")
p<-p+ geom_point(aes(y=PDI,col=log(Flowers)),size=7) + geom_line(aes(y=PDI)) 
p<-p+scale_fill_discrete(guide="none") + labs(y="",x="Date") + theme_bw() + scale_x_datetime(date_breaks = "2 months",date_labels = "%b %Y") + scale_color_continuous("Available Flowers",high='red',low='blue')
p

ggsave("Figures/NullNetwork.jpg",dpi=600,quality=100,pointsize=14,height=10,width=12)
ggsave("Figures/NullNetwork.svg",height=10,width=12)
```

## Difference from expected

```{r}
p<-ggplot(network.fl,aes(x=as.POSIXct(Date_C),fill=Metric)) + facet_wrap(~Metric,scales='free',ncol=5) + geom_linerange(aes(ymin=PDI-Lower,ymax=PDI-Upper),col="red")
p<-p+ geom_point(aes(y=PDI-Mean,col=log(Flowers)),size=7) + geom_line(aes(y=PDI-Mean)) 
p<-p+scale_fill_discrete(guide="none") + labs(y="",x="Date") + theme_bw() + scale_x_datetime(date_breaks = "2 months",date_labels = "%b %Y") + scale_color_continuous("Available Flowers",high='red',low='blue') 
p

```

##Resource Regression

```{r}
#Quick visualization 
p<-ggplot(network.fl[,],aes(log(Flowers),y=PDI-Mean,shape=as.factor(Year))) + geom_point(size=7)  + geom_smooth(aes(group=1),method="lm") + theme_bw() + labs(shape="Year",size="Interactions") + geom_text(aes(label=month.abb[as.numeric(Month)]),hjust=-0.25,size=4) + labs(x="Log Available Resources") + ylab("Score") 
p
```


```{r}
save.image("NetworkTime.RData")
```
