---
title: Appendix  B - Network Statistics.
author: "Ben Weinstein"
date: "2/1/2016"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
    keep_md: true
---

```{r,echo=F}
paste("Run Completed at",Sys.time())
```

```{r,echo=FALSE,warning=FALSE,message=FALSE}
require(knitr)
library(R2jags)
require(reshape2)
require(ggplot2)
library(scales)
library(dplyr)
library(stringr)
require(picante)
require(igraph)
require(bipartite)
library(chron)
library(gridExtra)
library(doSNOW)
library(foreach)
#Set Dropbox Location
#setwd to dropbox
droppath<-"C:/Users/Ben/Dropbox/"

#Set github path
gitpath<-"C:/Users/Ben/Documents/NetworkTime/"

setwd(gitpath)

opts_chunk$set(message=FALSE,warning=FALSE,fig.height=5,fig.width=10,echo=FALSE,cache=F,cache.path = 'NetworkTime_cache/', fig.path='figure/',fig.align = 'center')

```


************

```{r}
#load data if needed
#load("NetworkTime.Rdata")
```

#Read in data

From Generate.Rmd

```{r,fig.height=20,fig.width=13}

int<-read.csv("InputData/FinalObs.csv",row.names=1)

#just presence records
int<-int[int$State=="Used",]
```

#Overall Network

```{r}
tweb<-table(int$Hummingbird,int$Iplant_Double)
mtweb<-melt(tweb)
colnames(mtweb)<-c("Hummingbird","Iplant_Double","Visits")

#turn 0's to NA for plotting
mtweb$Visits[mtweb$Visits==0]<-NA

#get order of interactions
hord<-mtweb %>% group_by(Hummingbird) %>% filter(!is.na(Visits)) %>% summarize(n=n()) %>% arrange(n) %>% .$Hummingbird

mtweb$Hummingbird<-factor(mtweb$Hummingbird,levels=hord)

pord<-mtweb %>% group_by(Iplant_Double) %>% filter(!is.na(Visits)) %>% summarize(n=n()) %>% arrange(desc(n)) %>% .$Iplant_Double

mtweb$Iplant_Double<-factor(mtweb$Iplant_Double,levels=pord)

overall<-ggplot(mtweb,aes(y=Hummingbird,x=Iplant_Double,fill=Visits)) + geom_tile() + scale_fill_continuous(low="blue",high="red",na.value = "white") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Plant")
overall
```

##Calculate overall statistics.

```{r}
overallmetrics<-networklevel(as.matrix(tweb),level="lower",index=c("weighted connectance","weighted NODF"))
#melt into data frame
overallmetrics<-data.frame(Metric=names(overallmetrics),value=overallmetrics)
```

## Permutation tests

There is some discussion as to what the best null model for nestedness is. Should we constrain the matrix? Try both ways to see if it makes a difference in inference.

### Calculate randomized matrix

Maintaining the total number of observations, but shuffling where those visitations fall.

```{r}
null_all<-permatfull(tweb,times=500,fixedmar = "none")

null_metrics<-list()
for(x in 1:length(null_all$perm)){
  nm<-networklevel(null_all$perm[[x]],level="lower",index=c("weighted connectance","weighted NODF"))
  null_metrics[[x]]<-data.frame(Metric=names(nm),value=nm)
}

null_metrics<-melt(null_metrics)
ggplot(null_metrics,aes(x=value)) + geom_density(fill='grey40') + facet_wrap(~Metric,scales="free") + geom_vline(data=overallmetrics,aes(xintercept=value),col='red',linetype='dashed') + theme_bw()
```

```{r}
#Add overall network and measures underneath as a supp figure
num<-droplevels(null_metrics[null_metrics$Metric %in% c( "weighted connectance","weighted NODF"),])

ob<-droplevels(overallmetrics[overallmetrics$Metric %in% c("weighted connectance","weighted NODF"),])

levels(num$Metric)<-c("Connectance","Nestedness")
levels(ob$Metric)<-c("Connectance","Nestedness")

nulls<-ggplot(num,aes(x=value)) + geom_density(fill='grey40') + facet_wrap(~Metric,scales="free",nrow=1) + geom_vline(data=ob,aes(xintercept=value),col='red',linetype='dashed') + theme_bw()

jpeg("Figures/Overall.jpeg",height=6,width=11,units="in",res=400)
grid.arrange(overall,nulls,heights=c(.7,.3))
dev.off()
```

#Temporal change in network

##View Sub-Networks

```{r,fig.height=15}

int.M<-split(int,int$Availability,drop = T)

p<-lapply(int.M,function(x){

  tweb<-table(x$Hummingbird,x$Iplant_Double)
  mtweb<-melt(tweb)
  colnames(mtweb)<-c("Hummingbird","Iplant_Double","Visits")
  
  #turn 0's to NA for plotting
  mtweb$Visits[mtweb$Visits==0]<-NA
  mtweb<-mtweb[!is.na(mtweb$Visits),]
  
  #get order of interactions
  hord<-mtweb %>% group_by(Hummingbird) %>% filter(!is.na(Visits)) %>% summarize(n=n()) %>% arrange(n) %>% .$Hummingbird
  
  mtweb$Hummingbird<-factor(mtweb$Hummingbird,levels=hord)
  
  pord<-mtweb %>% group_by(Iplant_Double) %>% filter(!is.na(Visits)) %>% summarize(n=n()) %>% arrange(desc(n)) %>% .$Iplant_Double
  
  mtweb$Iplant_Double<-factor(mtweb$Iplant_Double,levels=pord)
  
  ggplot(mtweb,aes(y=Hummingbird,x=Iplant_Double,fill=Visits)) + theme_bw() +geom_tile() + scale_fill_continuous(low="blue",high="red",na.value = "white") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Plant") + ggtitle(unique(x$Availability)) 
})

grid.arrange(p[[1]],p[[2]])
```


## Null function

```{r}
nullC<-function(x){
  x<-droplevels(x)
  N<-x %>% group_by(ID) %>% n_groups()
  month.inter<-t(table(x$Hummingbird,x$Iplant_Double))

true_stat<-networklevel(month.inter,index=c("weighted connectance","weighted NODF"),level="lower")
  
true_stat<-melt(true_stat)

true_stat$variable<-rownames(true_stat)

# Create 500 random matrixes
r.int<-permatfull(month.inter,times=500,fixedmar="none")

#Create a functi\on that computes network metrics
null_total<-as.data.frame(
  t(sapply(r.int$perm,function(x){
  #Compute network statistics
    out<-networklevel(x,index=c("weighted connectance","weighted NODF"),level="lower")
    return(out)
    })))

#get the upper and lower 5th quantiles
null_total<-melt(null_total)
nullq<-group_by(null_total,variable) %>% summarize(Lower=quantile(value,0.05,na.rm=T),Upper=quantile(value,0.95,na.rm=T))

colnames(nullq)<-c("Metric","Lower","Upper")

#append true information
out<-merge(nullq,true_stat,by.x="Metric",by.y="variable")
colnames(out)[colnames(out) %in% "value"]<-"True"
out$N<-N
return(out)}
```

###Apply function to each period

```{r}

cl<-makeCluster(2,"SOCK")
registerDoSNOW(cl)
runs<-foreach(x=1:length(int.M),.packages=c("bipartite","reshape2","dplyr")) %dopar% {nullC(int.M[[x]])}
stopCluster(cl)

names(runs)<-names(int.M)
datHL<-melt(runs,id.vars=colnames(runs[[1]]))

colnames(datHL)[6]<-"HL"
```


#Observed versus null values of nestedness, connectance within time periods

```{r}
p<-ggplot(datHL,aes(x=HL,fill=Metric)) + facet_wrap(~Metric,scales='free',ncol=3) + geom_linerange(aes(ymin=Lower,ymax=Upper))
p<-p+ geom_point(aes(size=N,y=True)) + geom_line(aes(y=True)) 
p<-p+scale_fill_discrete(guide="none") + scale_size_continuous(range=c(2.5,5),"Observations") + labs(y="") + theme_bw()
print(p) 
ggsave("Figures/NullNetwork.jpg",dpi=600,quality=100,pointsize=14,height=10,width=12)
ggsave("Figures/NullNetwork.svg",height=10,width=12)

```

#Randomize with respect to overall resource membership.  

```{r}
rHL<-function(){
  
  #shuffle
  int$rHL<-sample(int$Availability)
  rint.M<-split(int,int$rHL,drop = T)
  
  r<-lapply(rint.M,function(x){
  month.inter<-t(table(x$Hummingbird,x$Iplant_Double))
  true_stat<-networklevel(month.inter,index=c("weighted connectance","weighted NODF"),level="lower")
  true_stat<-melt(true_stat)
  true_stat$variable<-rownames(true_stat)
  return(melt(true_stat))
  })
  r<-melt(r)
  colnames(r)[3]<-"Availability"
  return(r)
}

randomN<-list()
for(x in 1:500){
randomN[[x]]<-rHL()
  }

names(randomN)<-1:500
randomN<-melt(randomN,id.vars=colnames(randomN[[1]]))

colnames(randomN)<-c("Metric","value","HL","Iteration")

#find quantiles
randomQ<-randomN %>% group_by(Metric,HL) %>% summarize(Lower=quantile(value,0.05),Upper=quantile(value,0.95))

levels(datHL$Metric)<-c("Nestedness","Connectance")
randomQ$Metric<-as.factor(randomQ$Metric)
levels(randomQ$Metric)<-c("Connectance","Nestedness")

#compare to observed values. 
p<-ggplot(datHL,aes(x=HL,fill=Metric)) + facet_wrap(~Metric,scales='free',ncol=3)
p<-p+ geom_point(aes(size=N,y=True)) + geom_line(aes(y=True)) 
p<-p+scale_fill_discrete(guide="none") + scale_size_continuous(range=c(2.5,5),"Observations") + labs(y="") + theme_bw()
p + geom_linerange(data=randomQ,col='red',aes(ymin=Lower,ymax=Upper)) + xlab("Resource Availability")
ggsave("Figures/NetworkChange.jpg",dpi=300,height=2.5,width=5)
```

```{r}
save.image("NetworkTime.RData")
```
