---
title: "Observed Data"
author: "Ben Weinstein - Stony Brook University"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
    keep_md: true
  word_document: default
---

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(foreach)
library(doSNOW)
library(chron)
library(ggplot2)
library(knitr)
library(R2jags)
library(dplyr)
library(stringr)
library(gridExtra)
library(boot)
library(picante)
library(bipartite)

opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=4,echo=TRUE,cache=F,fig.align='center',fig.path="figureObserved/")

setwd("C:/Users/Ben/Documents/Occupy/")

set.seed(3)
#source functions

source("Functions.R")
```

```{r,echo=F,cache=FALSE}
paste("Run Completed at",Sys.time())
```

```{r}
#reload if needed
#load("Observed.Rdata")
```

#Observed dataset

```{r}
#read in flower morphology data, comes from Nectar.R
droppath<-"C:/Users/Ben/Dropbox/"
fl.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/FlowerMorphology.csv",sep=""))

#First row is empty
fl.morph<-fl.morph[-1,]

#Bring in Hummingbird Morphology Dataset, comes from
hum.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/HummingbirdMorphology.csv",sep=""))

#Bring in Interaction Matrix
int<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/Network/HummingbirdInteractions.csv",sep=""),row.names=1)

#one date error
int[int$DateP %in% '2013-07-25',"Month"]<-7

#Correct known taxonomic disagreements, atleast compared to traits
int[int$Iplant_Double=="Alloplectus purpureus","Iplant_Double"]<-"Glossoloma purpureum"
int[int$Iplant_Double=="Capanea affinis","Iplant_Double"]<-"Kohleria affinis"
int[int$Iplant_Double=="Columnea cinerea","Iplant_Double"]<-"Columnea mastersonii"
int[int$Iplant_Double=="Alloplectus teuscheri","Iplant_Double"]<-"Drymonia teuscheri"
int[int$Iplant_Double=="Drymonia collegarum","Iplant_Double"]<-"Alloplectus tetragonoides"

#Some reasonable level of presences, 25 points
keep<-names(which(table(int$Hummingbird) > 20))

int<-int[int$Hummingbird %in% keep,]

m.dat<-droplevels(int[colnames(int) %in% c("ID","Video","Time","Hummingbird","Sex","TransectID","Transect_R","Iplant_Double","Pierce","DateP","Month","ele","Type")])

#Does the data come from camera or transect?
m.dat$Type<-(is.na(m.dat$TransectID))*1

m.dat$Year<-years(as.Date(m.dat$DateP))
#one missing date
m.dat$Year[m.dat$Year %in% 2012]<-2013

#Number of bird species
h_species<-nlevels(m.dat$Hummingbird)

#Number of plant species
plant_species<-nlevels(m.dat$Iplant_Double)

#Get english name
dath<-merge(m.dat,hum.morph, by.x="Hummingbird",by.y="English",keep=all)

#Merge to flowers
int.FLlevels<-levels(factor(dath$Iplant_Double))

#Which flowers are we missing info for?
missingTraits<-int.FLlevels[!int.FLlevels %in% fl.morph$X]

#print(paste("Missing Trait Information:",missingTraits))
dath<-merge(dath,fl.morph, by.x="Iplant_Double",by.y="X")

#Drop piercing events, since they don't represent correlation
dath<-dath[!dath$Pierce %in% c("y","Y"),]
```

##Match Species to Morphology

```{r}
#remove species with less than  10 observations
keep<-names(which(table(dath$Hummingbird) > 20))

dath<-droplevels(dath[dath$Hummingbird %in% keep,])

#observed traitmatching
traitmatchF<-abs(t(sapply(hum.morph$Bill,function(x){x-fl.morph$TotalCorolla})))

rownames(traitmatchF)<-hum.morph$English
colnames(traitmatchF)<-fl.morph$Group.1
```

```{r}
#match names #Round to 2 decimals #Convert to cm for winbugs, avoids numerical underflow
traitmatchT<-round(traitmatchF[rownames(traitmatchF) %in% dath$Hummingbird,colnames(traitmatchF) %in% dath$Iplant_Double],2)/10

traitmatchT<-traitmatchT[sort(rownames(traitmatchT)),sort(colnames(traitmatchT))]
```

##Elevation ranges

Create a binary variable whether each observation was in a low elevation or high elevation transect. We have some species that just occur at the top of the gradient, and are not present in the sampling window of flowers at the low elevation.

Accounting for non-availability.
We have to figure out which plants were sampled in which periods, and if it was sampled, the non-detection are 0 if it wasn't the non-detection are NA. then remove all the Na's.

```{r}
elevH<-read.csv("InputData/HummingbirdElevation.csv",row.names=1)
head(elevH)
colnames(elevH)[5]<-"Elevation"
elevH$Bird<-1:nrow(elevH)

#high elevation or low elevation
elevP<-read.csv("InputData/PlantElevation.csv",row.names=1)
colnames(elevP)[5]<-"Elevation"
elevP$Plant<-1:nrow(elevP)
elevP$Iplant_Double<-as.character(elevP$Iplant_Double)

#Correct known taxonomic errors
elevP[elevP$Iplant_Double %in% "Alloplectus purpureus","Iplant_Double"]<-"Glossoloma purpureum"
elevP[elevP$Iplant_Double %in% "Capanea affinis","Iplant_Double"]<-"Kohleria affinis"
elevP[elevP$Iplant_Double %in% "Alloplectus teuscheri","Iplant_Double"]<-"Drymonia teuscheri"
elevP[elevP$Iplant_Double %in% "Columnea cinerea","Iplant_Double"]<-"Columnea mastersonii"
elevP[elevP$Iplant_Double %in% "Alloplectus tenuis","Iplant_Double"]<-"Drymonia tenuis"

#Merge to observed Data
#plants
dathp<-merge(dath,elevP,by="Iplant_Double")

#birds
datph<-merge(dathp,elevH,by="Hummingbird")

```

### Summarize Observations

```{r}
#ID for NA is holger transects, make the id's 1:n for each day of transect at each elevation, assuming no elevation was split across days.
datph$ID<-as.character(datph$ID)

noid<-datph[is.na(datph$ID),]

id_topaste<-paste(noid$Transect_R,noid$DateP,"Transect",sep="_")
datph[which(is.na(datph$ID)),"ID"]<-id_topaste

#Create year month combination
datph$Time<-paste(datph$Month,datph$Year,sep="_")

#Label survey type
datph$Survey_Type<-NA

mt<-!is.na(datph$TransectID)*1
datph$Survey_Type[mt==1]<-"Transect"
datph$Survey_Type[!datph$Survey_Type %in% "Transect"]<-"Camera"

indatraw<- datph %>% group_by(Hummingbird,Iplant_Double,Bird,Plant,Time,ID,DateP,Survey_Type) %>% summarize(Yobs=n(),Elev=mean(ele,na.rm=T),Transect_R=unique(Transect_R)) 

#add day ID
sdat<-split(indatraw,list(indatraw$ID),drop = T)

sdat<-lapply(sdat,function(x){
  x<-droplevels(x)
  x$Day<-as.numeric(as.factor(x$DateP))
  return(x)
})

indatraw<-rbind_all(sdat)

#Species names
for (x in 1:nrow(indatraw)){
  indatraw$Hummingbird[x]<-as.character(elevH[elevH$Bird %in% indatraw$Bird[x],"Hummingbird"])
  indatraw$Iplant_Double[x]<-as.character(elevP[elevP$Plant %in% indatraw$Plant[x],"Iplant_Double"])
}
```

What elevation transect is each observation in?
The camera data need to be inferred from the GPS point.

```{r}
#cut working best on data.frame
indatraw<-as.data.frame(indatraw)

#which elevation bin is each observation within
labs<-paste(seq(1300,2500,200),seq(1500,2700,200),sep="_")
indatraw$Transect_R[is.na(indatraw$Transect_R)]<-as.character(cut(indatraw[is.na(indatraw$Transect_R),]$Elev,seq(1300,2700,200),labels=labs))

#Elev for the transects is the midpoint
tran_elev<-indatraw[indatraw$Survey_Type=='Transect',"Transect_R"]
indatraw[indatraw$Survey_Type=='Transect',"Elev"]<-sapply(tran_elev,function(x){
  mean(as.numeric(str_split(x,"_")[[1]]))
})
```

```{r}
#match the traits
traitmelt<-melt(traitmatchT)
colnames(traitmelt)<-c("Hummingbird","Iplant_Double","Traitmatch")
```

##Absences - accounting for non-detection

We have more information than just the presences, given species elevation ranges, we have absences as well. Absences are birds that occur at the elevation of the plant sample, but were not recorded feeding on the flower.

```{r}
indatlong<-acast(indatraw,Bird~Plant~Time~ID~Day,value.var="Yobs")

indatlong[is.na(indatlong)]<-0
```

```{r}
#Only non-detections are real 0's, the rest are NA's and are removed.
#Plants not surveyed in that time period
#Hummingbirds not present at that elevation

for(t in 1:dim(indatlong)[3]){
  for(x in 1:dim(indatlong)[4]){
  
    #Remove non sampled plants 
  a<-indatlong[,,t,x,]

  #No observations at that plant
  toNA<-as.numeric(names(which(apply(a,2,sum)==0)))
  pres<-as.numeric(names(which(!apply(a,2,sum)==0)))
  
  if(length(pres)==0){next} else {
    #IDs are unique to time periods.

    indatlong[,,which(!dimnames(indatlong)[[3]] %in% dimnames(indatlong)[[3]][t]),x,]<-NA
  indatlong[,colnames(a) %in% toNA,t,x,]<-NA

  #Get elevation point of that sampling event
  cam<-dimnames(indatlong)[[4]][x]
  camelev<-indatraw %>% filter(ID==cam) %>% .$Elev %>% unique()
      
  #for each hummingbird, was that camera within elevation?
  for(i in 1:dim(a)[1]){
      low<-elevH[elevH$Bird == rownames(a)[i],"Low"]
      high<-elevH[elevH$Bird == rownames(a)[i],"High"]
        
        #if not in elev range, set to NA
        
        if(!((low < camelev) & (camelev < high))){
          #if you had a wandering individual outside range, allow interaction to occur.
                indatlong[i,,t,x,]<-NA
        }
      }
    }
  }
}
### There can't be absences in days that weren't sampled.
for (x in 1:dim(indatlong)[4]){
  cam<-indatlong[,,,x,]
  for (y in 1:dim(cam)[4]){
    sc<-sum(cam[,,,y],na.rm=T)
    if (sc ==0){
      indatlong[,,,x,y]<-NA
    }
  }
}

#melt and remove Na's
indat<-melt(indatlong)
indat<-indat[!is.na(indat$value),]

colnames(indat)<-c("Bird","Plant","Time","ID","Day","Yobs")
```

```{r}
#remerge the time period data
Timelookup<-indatraw %>% dplyr::select(ID,Transect_R,Survey_Type) %>% group_by(ID,Transect_R,Survey_Type) %>% distinct() %>% arrange(ID)

#Get time information
indat<-merge(indat,Timelookup,by=c("ID"))

#Species names
for (x in 1:nrow(indat)){
  indat$Hummingbird[x]<-as.character(elevH[elevH$Bird %in% indat$Bird[x],"Hummingbird"])
  indat$Iplant_Double[x]<-as.character(elevP[elevP$Plant %in% indat$Plant[x],"Iplant_Double"])
}

#Get trait information
#match the traits
indat<-merge(indat,traitmelt,by=c("Hummingbird","Iplant_Double"))
```

Reformat index for jags.
Jags needs a vector of input species 1:n with no breaks.

```{r}
#Easiest to work with jags as numeric ordinal values
indat$Hummingbird<-as.factor(indat$Hummingbird)
indat$Iplant_Double<-as.factor(indat$Iplant_Double)
indat$jBird<-as.numeric(indat$Hummingbird)
indat$jPlant<-as.numeric(indat$Iplant_Double)

jagsIndexBird<-data.frame(Hummingbird=levels(indat$Hummingbird),jBird=1:length(levels(indat$Hummingbird)))
 
jagsIndexPlants<-data.frame(Iplant_Double=levels(indat$Iplant_Double),jPlant=1:length(levels(indat$Iplant_Double)))

#Similiarly, the trait matrix needs to reflect this indexing.
jTraitmatch<-traitmatchT[rownames(traitmatchT) %in% unique(indat$Hummingbird),colnames(traitmatchT) %in% unique(indat$Iplant_Double)]
```

#Resources at each month

In our model the covariate is indexed at the scale at which the latent count is considered fixed. This means we need the resource availability per month across the entire elevation gradient for each point.

```{r}
fl<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/FlowerTransects/FlowerTransectClean.csv",sep=""))

#select desired columns
fltransects<-fl %>% dplyr::select(Iplant_Double,month,Year,ID,Transect.ID,Family,Genus,Species,Height,Total_Flowers,Transect_R,Date,lon,lat,ele)

#dates need a bit of formatting
fltransects$DateP<-as.POSIXlt(fltransects$Date,format="%d/%m/%y")

#which are NA needs to date/month reversed. ugly.
fltransects[which(is.na(fltransects$DateP)),"DateP"]<-as.character(as.POSIXlt(as.character(fltransects[which(is.na(fltransects$DateP)),"Date"]),format="%m/%d/%Y"))

#remove without elevation records (less than 1% of data)
fltransects<-fltransects[!is.na(fltransects$ele),]

#month should be capital 
colnames(fltransects)[colnames(fltransects) %in% "month"]<-"Month"

#Count of flowers per month
fltransects$DateP<-as.character(fltransects$DateP)
flower.month<-group_by(fltransects,Month,Year,Transect_R,DateP) %>% dplyr::summarise(Flowers=sum(Total_Flowers,na.rm=TRUE))  %>% filter(Flowers>20)
  
#Make month abbreviation column, with the right order
flower.month$Month.a<-factor(month.abb[flower.month$Month],month.abb[c(1:12)])

#Make year factor column
flower.month$Year<-as.factor(flower.month$Year)

#get quantile for each transect
thresh<-melt(group_by(flower.month) %>% summarize(Threshold=mean(log(Flowers))))
             
flower.month<-merge(flower.month,thresh)
flower.month$Availability<-as.factor((log(flower.month$Flowers)>flower.month$value) * 1)

#plot
ggplot(flower.month,aes(col=Availability,x=Month.a,log(Flowers),shape=as.factor(Year))) + geom_point(size=3) + theme_bw()  + geom_smooth(aes(group=1)) + ylab("Flowers") + xlab("Month") + facet_wrap(~Transect_R,scales="free_y") + labs(shape="Year", y= "Log Available Flowers") + scale_x_discrete(breaks=month.abb[seq(1,12,2)]) + scale_color_manual(labels=c("Low","High"),values=c("black","red")) 

#turn min and max elvation into seperate columns for the range
flower.month$minElev<-as.numeric(str_extract(flower.month$Transect_R,"(\\d+)"))
flower.month$maxElev<-as.numeric(str_match(flower.month$Transect_R,"(\\d+)_(\\d+)")[,3])
```

```{r}
#summarize per month
#Count of flowers per month
Fmonth<-flower.month %>%group_by(Month,Year,Transect_R) %>% summarise(Flowers=mean(Flowers)) %>% summarize(Flowers=mean(Flowers)) %>% mutate(Time=paste(Month,Year,sep="_")) 

#Binary model
Fmonth$Availability<-(log(Fmonth$Flowers) > mean(log(Fmonth$Flowers)) )*1 
ggplot(Fmonth,aes(x=as.factor(Month),y=log(Flowers))) + geom_point(size=4,aes(col=as.factor(Availability))) + geom_line(aes(group=Year)) + labs(x="Month",y="Log Transformed Flower Count")
indat<-merge(indat,Fmonth,by="Time")

indat<-dcast(indat,...~Survey_Type,value.var="Yobs",fill=0)

#Turn Time and ID into numeric indexes
indat$jTime<-as.numeric(as.factor(indat$Time))
indat$jID<-as.numeric(as.factor(indat$ID))
```


#Hierarchical Occupancy Model

For hummingbird species i feeding on plant species j observed at time k and sampling event d. 

$$ YTransect_{i,j,k,d} \sim B(N_{i,j,k},\omega_{Transect}) $$

$$ YCamera_{i,j,k,d} \sim B(N_{i,j,k},\omega_{Camera}) $$
$$ \omega_{Camera} <- \phi_{Camera} * EffortCamera_k $$
$$ \omega_{Transect} <- \phi_{Transect}* EffortTransect_k $$
$$ N_{i,j,k} \sim Pois(\lambda_{i,j,k}) $$
$$ log(\lambda_{i,j,k}) = \alpha_i + \beta_{1,i} * Traitmatch_{i,j} + \beta_{2,i} *Resources_k + \beta_{3,i} * Traitmatch_{i,j} * Resources_k $$

**Priors**

$$ \phi_{Camera} \sim U(0,1) $$
$$ \phi_{Transect} \sim U(0,1) $$
$$\alpha_i \sim N(intercept,\tau_{\alpha})$$
$$\beta_{1,i} \sim N(\gamma_1i,\tau_{\beta_1})$$
$$\beta_{2,i} \sim N(\gamma_2i,\tau_{\beta_2})$$
$$\beta_{3,i} \sim N(\gamma_3i,\tau_{\beta_3})$$

**Hyperpriors**

Group Level Means

$$\gamma_{1,i} \sim N(0,0.0001)$$
$$\gamma_{2,i} \sim N(0,0.0001)$$
$$\gamma_{3,i} \sim N(0,0.0001)$$
$$ intercept \sim N(0,0.0001)$$

Group Level Variance

$$\tau_{\alpha} \sim Gamma(0.0001,0.0001)$$
$$\tau_\beta1 \sim Gamma(0.0001,0.0001)$$
$$\tau_\beta2 \sim Gamma(0.0001,0.0001)$$
$$\tau_\beta3 \sim Gamma(0.0001,0.0001)$$

**Derived quantities**

$$\sigma_{intercept} = \sqrt[2]{\frac{1}{\tau_\alpha}}$$
$$\sigma_{slope1} = \sqrt[2]{\frac{1}{\tau_{\beta_1}}}$$
$$\sigma_{slope2} = \sqrt[2]{\frac{1}{\tau_{\beta_2}}}$$
$$\sigma_{slope3} = \sqrt[2]{\frac{1}{\tau_{\beta_3}}}$$

```{r,eval=F,strip.white=T}

#Source model
source("Bayesian/NmixturePoissonRagged2m.R")

#print model
writeLines(readLines("Bayesian/NmixturePoissonRagged2m.R"))

#Input Data
Dat <- c('Yobs_camera','Yobs_transect','Birds','Bird','Plant','Time','Plants','Times','resources','Nobs','cam_surveys','trans_surveys','Traitmatch')

#Inits
InitStage <- function(){
  #A blank Y matrix - all present
  initY<-array(dim=c(Birds,Plants,Times),data=max(Yobs_transect,na.rm=T))
  initB<-as.numeric(matrix(nrow=Birds,ncol=1,data=.1))

list(beta1=initB,beta2=initB,beta3=initB,alpha=rep(.5,Birds),intercept=0,tau_alpha=0.1,tau_beta1=0.1,tau_beta2=0.1,tau_beta3=0.1,gamma1=0,gamma2=0,gamma3=0,dtrans=0.5,dcam=0.5,N=initY)}

#Parameters to track
ParsStage <- c("alpha","beta1","beta2","beta3","intercept","sigma_int","sigma_slope1","sigma_slope2","sigma_slope3","gamma1","gamma2","gamma3","dtrans","dcam")

#MCMC options

ni <- 200  # number of draws from the posterior
nt <- max(c(1,ni*.0001))  #thinning rate
nb <- ni*.85 # number to discard for burn-in
nc <- 2  # number of chains

#Jags

  Yobs_camera = indat$Camera
  Yobs_transect = indat$Transect
  Birds=max(indat$jBird)
  Bird=indat$jBird
  Plant=indat$jPlant
  Time=indat$jTime
  Plants=max(indat$jPlant)
  Times=max(indat$jTime)
  resources=TimeResources
  Nobs=nrow(indat)
  cam_surveys=(indat$Survey_Type=="Camera")*1
  trans_surveys=(indat$Survey_Type=="Transect")*1
  Traitmatch=Traitmatch

  m2<-do.call(jags.parallel,list(Dat,InitStage,ParsStage,model.file="Bayesian/NmixturePoissonRagged2m.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc))
```

```{r,eval=F}
#recompile if needed
load.module("dic")
runs<-40000
recompile(m2)
m2<-update(m2,n.iter=runs,n.burnin=runs*.9,n.thin=3)
```

```{r}
#extract par to data.frame
pars_detect<-extract_par(m2,data=indat,Bird="jBird",Plant="jPlant")

#name
pars_detect$Model<-"Occupancy"
```

##Assess Convergence

```{r,cache=FALSE,fig.width=13,fig.height=5}
###Chains
ggplot(pars_detect[pars_detect$par %in% c("detect","alpha","beta"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")
```

###Hierarcichal Posteriors

```{r,fig.height=5,fig.width=11}
ggplot(pars_detect[pars_detect$par %in% c("gamma","intercept","sigma_int","sigma_slope"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
ggsave("Figures/BothObs.svg",height=5,width=7)
ggsave("Figures/BothObs.jpg",height=5,width=7,dpi=300)
```

```{r}
#Bind together the two models
parsObs<-rbind(pars_detect,pars_dniave)
```

##Posteriors

```{r,cache=FALSE,fig.width=11,fig.height=14}
###Posterior Distributions
ggplot(parsObs[parsObs$par %in% c("detect","alpha","beta"),],aes(x=estimate,fill=Model)) + geom_histogram(position='identity') + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Detection Probability")
```

```{r,cache=FALSE,fig.width=10,fig.height=5}
#Detection figure
ggplot(parsObs[parsObs$par %in% c("detect"),],aes(x=as.factor(species),y=estimate,fill=Model)) + geom_violin() + ggtitle("Estimate of parameters") + theme_bw() + ggtitle("Detection Probability") +facet_wrap(~Model,scales="free") 

pars_detect<-merge(pars_detect,jagsIndexBird,by.x="species",by.y="jBird",all.x=T)

ggplot(pars_detect[pars_detect$par %in% c("detect"),],aes(x=estimate)) + geom_histogram() + ggtitle("Posterior Distribution") + theme_bw() + facet_wrap(~Hummingbird,ncol=5) + xlab("Probability of Detection")
ggsave("Figures/DetectionProb.jpg",dpi=300,height=7,width=11)
```

```{r,cache=FALSE,fig.height=5,fig.width=13}
ggplot(parsObs[parsObs$par %in% c("gamma","intercept","sigma_int","sigma_slope"),],aes(x=estimate,fill=Model)) + geom_histogram() + ggtitle("Trait matching regression parameters") + facet_wrap(~par,scale="free",nrow=2) + theme_bw() 
```

##Species Predictions

```{r,fig.height=10,fig.width=11}
castdf<-dcast(parsObs[parsObs$par %in% c("beta","alpha"),], species +Chain +Model+ Draw~par,value.var="estimate")

#Turn to species level
castdf$species<-factor(castdf$species,levels=1:max(as.numeric(castdf$species)))

species.split<-split(castdf,list(castdf$species,castdf$Model))

species.traj<-list()

for(d in 1:length(species.split)){
  x<-species.split[[d]]
  #species name
  index<-jagsIndexBird[unique(x$species),"Hummingbird"]
  
  #range of trait distances
  tsp<-indat %>% filter(Hummingbird==index) %>% distinct(Traitmatch) %>% .$Traitmatch
  species.traj[[d]]<-trajF(alpha=x$alpha,beta=x$beta,x=tsp)
  }

names(species.traj)<-names(species.split)

species.traj<-melt(species.traj,id.var=colnames(species.traj[[1]]))

#split out names and model
species.traj[,c("Index","Model")]<-colsplit(species.traj$L1,"\\.",c("Index","Model"))

spe<-merge(species.traj,jagsIndexBird,by.x="Index",by.y="jBird")

#match colnames

#plot and compare to original data
ggplot(data=spe[,],aes(x=x)) + geom_point(data=indat,aes(x=Traitmatch,y=Yobs)) + geom_ribbon(aes(ymin=lower,ymax=upper,fill=Model),alpha=0.2)  + geom_line(aes(y=mean,col=Model),size=1) + theme_bw() + ylab("Interactions") + xlab("Difference between Bill and Corolla Length") + facet_wrap(~Hummingbird,scales="free",ncol=3)+ labs(fill="Model")  + ylab("Interactions per day")

ggsave("Figures/SpeciesPredictionsBoth.jpg",dpi=300,height=8,width=10)
```

###Overall predicted relationship 

Does accounting for non-independence and detection change our estimate of trait matching?

```{r,fig.height=6,fig.width=10}

castdf<-dcast(parsObs[parsObs$par %in% c("gamma","intercept"),], Model+Chain + Draw~par,value.var="estimate")

castdf<-split(castdf,castdf$Model)
#calculated predicted y
predy<-rbind_all(lapply(castdf,function(i){
  #calculate trajectory and append model
  pr<-trajF(alpha=i$intercept,beta=i$gamma,x=indat$Traitmatch)  
  pr$Model<-unique(i$Model)
  return(pr)
  }))

#plot and compare to original data
ggplot(data=predy,aes(x=x,col=Model,fill=Model)) + geom_point(data=indat,aes(x=Traitmatch,y=Yobs),col="black",fill="black",size=2) + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.6)  + geom_line(aes(y=mean),size=1) + theme_bw() + ylab("Interactions") + xlab("Difference between Bill and Corolla Length") + ylab("Interactions per day") + ylim(0,17)
ggsave("Figures/ObsResults.jpg",dpi=300,height=5,width=7)
```

##Discrepancy 

The goodness of fit is a measured as chi-squared. The expected value for each day is the detection rate * the estimate intensity of interactions. The expected value is compared to the observed value of the actual data. In addition, a replicate dataset is generated from the posterior predicted intensity. Better fitting models will have lower discrepancy values and be 
Better fitting models are smaller values and closer to the 1:1 line. A perfect model would be 0 discrepancy. This is unrealsitic given the stochasticity in the sampling processes. Rather, its better to focus on relative discrepancy. In addition, a model with 0 discrepancy would likely be seriously overfit and have little to no predictive power.

```{r,fig.height=4,fig.width=8,message=F,warning=F}
fitstat<-parsObs[parsObs$par %in% c("fit","fitnew"),]
fitstat<-dcast(fitstat,Model+Draw+Chain~par,value.var="estimate")

ymin<-round(min(fitstat$fit))
ymax<-round(max(fitstat$fit))
ab<-data.frame(x=0:ymax,y=0:ymax)
disc_obs<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data",col="Model")  + ggtitle("Empirical Data") + geom_line(data=ab,aes(x=x,y=y)) + coord_fixed() + ylim(ymin=0,ymax=max(max(c(fitstat$fit,fitstat$fitnew)))) + xlim(xmin=0,xmax=max(max(c(fitstat$fit,fitstat$fitnew))))
disc_obs
ggsave("Figures/ObservedDiscrepancy.jpeg",width = 5,height=10)
```

##Detection table

```{r}
dp<-group_by(pars_detect[pars_detect$par %in% c("detect"),],species) %>% summarise(mean=round(mean(estimate,na.rm=T),3)*100,lower=round(quantile(estimate,0.025,na.rm=T),3)*100,upper=round(quantile(estimate,0.975,na.rm=T),3)*100)

tab<-merge(dp,jagsIndexBird,by.x="species",by.y="jBird")[,-1]
tab[,c(4,1,2,3)]
write.csv(tab[,c(4,1,2,3)],"Figures/Table1.csv")
```

##Sampling intensity and detection for each hummingbird species

The probability of missing a species at each daily visit is 1 - detection probability.

The probability of missing a species on sequential visits is (1- detection probability) * (1 - detection probability).

We are interested in the number of sampling events that minimize this value to some reasonable threshold. I have chosen 0.05 by convention. 

The following figure represesent the estimated number of daily surveys to capture a hummingbird event given that we know it occurs. These data can be thought of as successful draws from a negative binomial distribution.

It is easiest to interpret as the number of days until you are likely to see an interaction, so i prefer to calculate:

$$ 1-(1-p)^n $$
$$ p = probability of detection$$
$$ n = days sample $$

```{r,fig.width=6,fig.height=8}
dp<-function(n,p){
  1-((1-p)^n)
}

ts<-split(tab,tab$Hummingbird,drop=T)
detectd<-lapply(ts,function(x){
  meanD<-dp(n=1:10,p=x$mean/100)
  lowerD<-dp(n=1:10,p=x$lower/100)
  upperD<- dp(n=1:10,p=x$upper/100)
  data.frame(Days=1:10,mean=meanD,lower=lowerD,upper=upperD)
})

md<-melt(detectd,id.var="Days")
md<-dcast(md,...~variable)

#get the 0.5 line
dpn<-function(t,p){
  n<-(1 - (1-t))/(p/100)
  return(n)
}

#for each bird get the upper and middle estimate for 50% chance.
daydf<-list()
for (x in 1:nrow(tab)){
  mean_day=dpn(t=0.5,tab$mean[x])
  lower_day=dpn(t=0.5,tab$lower[x])
  upper_day=dpn(t=0.5,tab$upper[x])
  daydf[[x]]<-data.frame(L1=tab$Hummingbird[x],mean=mean_day,lower=lower_day,upper=upper_day)
}
daydf<-rbind_all(daydf)

ggplot(md,aes(x=Days,fill=L1,y=mean,ymin=lower,ymax=upper)) + geom_ribbon(alpha=.5) + geom_line() + facet_wrap(~L1,nrow=2,scale="free_x")  + ylab("Probability of detecting a interaction") + scale_fill_discrete(guide="none") + theme_bw() + scale_x_continuous(breaks=seq(0,10,2),limits=c(0,10))+ geom_rect(fill='grey',data=daydf,alpha=0.4,aes(xmax=upper,xmin=lower,ymin=0,ymax=Inf)) + ylim(0,1)
ggsave("Figures/DetectionDays.jpeg",height=5,width=10,dpi=300) 
```

The number of days it would take to have 50% confidence you have sampled enough to capture known interactions is the x axis value where the dotted line hits the curve.

Number of samples per species

##DIC Table

```{r}
m2_niave$BUGSoutput$DIC
m2$BUGSoutput$DIC
```

#Predicted versus Observed Data

```{r}
m<-max(traitmatchT)-traitmatchT

mat<-indat %>% group_by(jBird,jPlant) %>% summarize(n=sum(Yobs))
true_state<-acast(mat,jBird~jPlant,fill=0)

#get species names?

#trait liklihood
dmultinom(true_state,prob=m,log=T)
paste("Correlation coefficient is:", round(cor(c(true_state),c(m),method="spearman"),2))

```

###Test Statistic
Chisquared statistic is (observed-expected)^2/(expected + 0.5), we add the 0.5 to avoid dividing by 0. For each combination of birds and plants, predicted versus  observed for each data point.

##Multinomial prediction

Based on trait differences.

```{r}

#define discrep function
chisq<-function(o,e){(o-e)^2/(e+0.5)}

nullm<-function(){
r<-mgen(m,sum(true_state),keep.species = F)
}

#same number of draws as bayesian
cl<-makeCluster(20,"SOCK")
registerDoSNOW(cl)
mult_matrix<-foreach(x=1:length(pars_detect[pars_detect$par %in% "fit","estimate"]),.packages=c("bipartite","reshape2")) %dopar% nullm()
stopCluster(cl)

mats<-lapply(mult_matrix,function(r){
rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
colnames(rmerge)<-colnames(true_state)
  rownames(rmerge)<-rownames(true_state)
#for each position what is the chisq
for (x in 1:nrow(r)){
  for (y in 1:ncol(r)){
   rmerge[x,y]<-chisq(o=r[x,y],e=true_state[x,y])
  }
}

#sum of chisq driscrepancy
return(rmerge)
})

multi_disc<-sapply(mats,function(x) sum(x))

qplot(multi_disc)+ xlab("Chisquared Discrepancy for Multimonial Liklihood") + geom_vline(xintercept=mean(multi_disc),col='red',linetype='dashed')
```

#Compare Bayesian Occupancy and Multinomial using true known interactions

## Poisson GLMM

```{r}
N_niave<-pars_dniave[ pars_dniave$par %in% "ynew",]

bydraw<-split(N_niave,list(N_niave$Chain,N_niave$Draw))

occ_nodetect_matrix<-lapply(bydraw,function(x){
  r<-acast(x,species ~ plant,value.var = "estimate",fun.aggregate = sum)
})

#calculate discrep on those deviates
occ_nodetect<-lapply(occ_nodetect_matrix,function(r){
    #for each position what is the chisq
  rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
  colnames(rmerge)<-colnames(true_state)
  rownames(rmerge)<-rownames(true_state)
  for (x in 1:nrow(r)){
    for (y in 1:ncol(r)){
     rmerge[x,y]<-chisq(o=r[x,y],e=true_state[x,y])
      }
    }
  return(rmerge)
})

names(occ_nodetect)<-1:length(occ_nodetect)
names(occ_nodetect_matrix)<-1:length(occ_nodetect_matrix)
```

##With Detection

```{r}
N<-pars_detect[pars_detect$par %in% "ynew",]
bydraw<-split(N,list(N$Chain,N$Draw))
occ_matrix<-lapply(bydraw,function(x){
  r<-acast(x,species ~ plant,value.var = "estimate",fun.aggregate = sum)
  })

#calculate discrep for those aggregated matrices
occ<-lapply(occ_matrix,function(r){
    #for each position what is the chisq
  rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
  colnames(rmerge)<-colnames(true_state)
  rownames(rmerge)<-rownames(true_state)
  for (x in 1:nrow(r)){
    for (y in 1:ncol(r)){
     rmerge[x,y]<-chisq(o=r[x,y],e=true_state[x,y])
      }
  }
  return(rmerge)
})

names(occ)<-1:length(occ)
names(occ_matrix)<-1:length(occ_matrix)

```

##Compare to observed data

```{r,fig.width=12}
mmat<-melt(true_state)
colnames(mmat)<-c("Bird","Plant","True_State")

#append to predicted matrices

#occupancy with detection
mocc<-melt(occ_matrix)
colnames(mocc)<-c("Bird","Plant","Occupancy","Iteration")
simdat<-merge(mocc,mmat,by=c("Bird","Plant"))

#occupancy with nodetection
moccd<-melt(occ_nodetect_matrix)
colnames(moccd)<-c("Bird","Plant","Poisson GLMM","Iteration")

simdat<-merge(simdat,moccd,by=c("Bird","Plant","Iteration"))

#multinomial
multmats<-melt(mult_matrix)
colnames(multmats)<-c("Bird","Plant","Multinomial","Iteration")
simdat<-merge(simdat,multmats,by=c("Bird","Plant","Iteration"))


simdat<-melt(simdat,measure.vars = c("Occupancy","Poisson GLMM","Multinomial"))

ggplot(simdat,aes(x=True_State,y=value,col=variable)) + geom_point() + geom_abline() + labs(col="Model") + ylab("Predicted State") + xlab("True State") + theme_bw() + facet_wrap(~variable)

```

##Summary of discrepancy of predicted matrices

```{r}
#Multinomial
multi_disc<-sapply(mats,function(x) mean(x))

#occupancy without detection
occno_disc<-sapply(occ_nodetect,function(x) mean(x))

#Occupancy with detection
occ_disc<-sapply(occ,function(x) mean(x))

#compared to bayesian
ggplot(data.frame(multi_disc)) + geom_histogram(aes(x=multi_disc),fill="blue",alpha=.6)+ xlab("Chi-squared Discrepancy") + geom_histogram(data=data.frame(occ_disc),aes(x=occ_disc),fill="red",alpha=.6) + theme_bw() +geom_vline(aes(xintercept=mean(occ_disc)),linetype="dashed",col="red")+ geom_vline(xintercept=mean(multi_disc),linetype="dashed",col="blue") + geom_histogram(data=data.frame(occno_disc),aes(x=occno_disc),fill="orange",alpha=.6) + geom_vline(aes(xintercept=mean(occno_disc)),linetype="dashed",col="orange")
```

##Comparison of summary statistics for all three approaches

```{r}
d<-list(Occupancy=occ,Multinomial=mats,Poisson_GLM=occ_nodetect)
d<-melt(d)
colnames(d)<-c("Bird","Plant","value","Iteration","Model")

d %>% group_by(Model,Iteration) %>% summarize(mean=mean(value),sd=sd(value),sum=sum(value)) %>% group_by(Model) %>% summarize(mean_mean=round(mean(mean),2),mean_sd=round(sd(mean),2),mean_sum=round(mean(sum),2))
```

Merge with morphological data.

```{r}
jT<-indat %>% group_by(Bird=jBird,Plant=jPlant) %>% summarize(Traitmatch=unique(Traitmatch))
simdat<-merge(simdat,jT,by=c("Bird","Plant"))

mmat<-merge(mmat,jT,by=c("Bird","Plant"))
```

#Predicted total number of visits based on morphology

```{r}
simT<-simdat %>% group_by(variable,Traitmatch) %>% summarize(Lower=quantile(value,0.05),Upper=quantile(value,0.95),y=mean(value))

ggplot(simT,aes(x=Traitmatch)) + geom_ribbon(aes(ymin=Lower,ymax=Upper,fill=variable),alpha=0.6) + geom_line(aes(y=y,col=variable),linetype='dashed') + theme_bw() + facet_wrap(~variable,nrow=3) + geom_point(data=mmat,aes(y=True_State))
```


```{r}
gc()
save.image("Observed.Rdata")
```


