---
title: "Trait-matching and Available Resources: Randomizations"
author: "Ben Weinstein - Stony Brook University"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
    keep_md: true
  word_document: default
---

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(foreach)
library(doSNOW)
library(chron)
library(ggplot2)
library(knitr)
library(R2jags)
library(dplyr)
library(stringr)
library(gridExtra)
library(boot)
library(picante)
library(bipartite)
library(VennDiagram)

opts_chunk$set(message=FALSE,warning=FALSE,fig.width=9,fig.height=7,echo=TRUE,cache=F,fig.align='center',fig.path="figureObserved/")

#source functions
```

```{r,echo=F,cache=FALSE}
paste("Run Completed at",Sys.time())
```

```{r}
#reload if needed
#load("Randomization.Rdata")
```

#Load in data

```{r}
#From Generate.Rmd - Interaction Data
int<-read.csv("InputData/FinalObs.csv")
int$State<-(int$Yobs==1)*1
int$State<-as.factor(int$State)
levels(int$State)<-c("Unused","Used")
```

```{r}
#Morphology
fl.morph<-read.csv("InputData/FlowerMorphology.csv")

#First row is empty
fl.morph<-fl.morph[-1,]

#Bring in Hummingbird Morphology Dataset, comes from
hum.morph<-read.csv("InputData/HummingbirdMorphology.csv")

#taxonomy change, we are calling them Crowned Woodnymph's now.
hum.morph$English<-as.character(hum.morph$English)
hum.morph$English[hum.morph$English %in% "Green-crowned Woodnymph"]<-"Crowned Woodnymph"
```

##Randomization 1
Is the observed distribution of choosen corollas different than the background distribution of availabile corollas.

This is most common definition of specialization.

# Across time

```{r}

#randomization function
rand1<-function(x,r){
  
#Calculate mean function
nullc<-function(x){
    x$State<-sample(x$State)
    null_state<- x %>% filter(State=="Used") %>% group_by(Hummingbird) %>%   summarize(mean=mean(Traitmatch))
      return(null_state)
  }

  cl<-makeCluster(2,"SOCK")
  registerDoSNOW(cl)
  nullframe<-foreach(j=1:r,.packages="dplyr") %dopar% {nullc(x)}
  stopCluster(cl)
  
  names(nullframe)<-1:r
  nullframe<-melt(nullframe,id.vars=c("Hummingbird"))
}
```

```{r}
#True state
true_state<- int %>% filter(State=="Used") %>% group_by(Hummingbird) %>% summarize(mean=mean(Traitmatch),Lower=quantile(Traitmatch,0.05),Higher=quantile(Traitmatch,0.95))
mtruestate<-melt(true_state,measure.vars=c("mean","Higher","Lower"))

nullframe<-rand1(int,100)
```

```{r}
ggplot(nullframe,aes(y=value,x=Hummingbird)) + geom_violin(fill="black",alpha=.3) + geom_point(data=true_state,aes(y=mean,x=Hummingbird),col="red",size=4) + theme_bw() + coord_flip() + ylim(0,max(true_state$mean)) + ylab("Difference between Bill and Corolla Length")
```

Are those true distributions within reasonable confidence bands?

```{r}
sdat<-split(nullframe,nullframe$variable)
tstate<-sapply(sdat,function(d){
  sp<-split(d,d$Hummingbird,drop = T)
  l<-sapply(sp,function(y){
    #true state
    ts<-true_state[true_state$Hummingbird %in% unique(y$Hummingbird),as.character(unique(d$variable))]
    f<-ecdf(y$value)
    f(ts[[1]])
  })
})

#melt for plotting
mtstate<-melt(tstate)

ord<-hum.morph %>% arrange(desc(Bill)) %>% select(English) %>% .$English

mtstate$Var1<-factor(mtstate$Var1,levels=ord)

p<-ggplot(mtstate,aes(x=Var1,y=Var2,fill=(value<0.05 | value > 0.95))) + geom_tile() + coord_flip() + labs(x="Hummingbird",y="Metric",fill="Difference between\nObserved and Null Distribution\n(alpha=0.05)") + ggtitle("Used versus Background")
p
```

##Change through time

```{r}
#Create time stamp
int$TimeStamp<-as.POSIXct(int$DateP)

#Set window size
cutm<-"1 month"
int$Date_C<-cut(int$TimeStamp,breaks=cutm)
int.M<-split(int,int$Date_C,drop = T)

#True state for each species

truemonths<-lapply(int.M,function(x){
  #True state
  true_state<- x %>% filter(State=="Used") %>% group_by(Hummingbird) %>% summarize(mean=mean(Traitmatch),Lower=quantile(Traitmatch,0.05),Higher=quantile(Traitmatch,0.95))
  mtruestate<-melt(true_state,measure.vars=c("mean","Higher","Lower"))
  true_state$Date_C<-unique(x$Date_C)
  return(true_state)
})

#bind together
truemonths<-rbind_all(truemonths)
  
truemonths$Date_C<-as.POSIXct(truemonths$Date_C,tz = "")
```

Empirical matching

```{r}

int$month<-months(as.POSIXct(int$Date_C))
int$month<-as.factor(int$month)
int$month<-factor(int$month,levels=month.name)
int$year<-years(as.POSIXct(int$Date_C))

ggplot(int[int$State=="Used" & int$Hummingbird =="Gorgeted Sunangel",],aes(x=month)) + geom_violin(aes(y=Traitmatch,fill=year),position=position_dodge(width=0.5)) + theme_bw() + ylab("Corolla Length Selected") + facet_wrap(~Hummingbird,scales="free",nrow=5) + scale_x_discrete(breaks=month.name[seq(1,12,2)]) + labs(fill="Year")

emp<-int %>% group_by(Date_C,Hummingbird) %>% filter(State=="Used") %>% summarize(mean=mean(Traitmatch),Upper=quantile(Traitmatch,0.95),Lower=quantile(Traitmatch,0.05))

emp$month<-months(as.POSIXct(emp$Date_C))
emp$month<-as.factor(emp$month)
emp$month<-factor(emp$month,levels=month.name)
emp$year<-years(as.POSIXct(emp$Date_C))

ggplot(emp[emp$Hummingbird=="Tawny-bellied Hermit",],aes(x=month)) + geom_point(aes(y=mean,col=year),position=position_dodge(width=0.5)) + theme_bw() + ylab("Corolla Length Selected") + facet_wrap(~Hummingbird,scales="free",nrow=5)  + geom_linerange(aes(ymin=Lower,ymax=Upper,col=year),position=position_dodge(width=0.5)) + scale_x_discrete(breaks=month.name[seq(1,12,2)])
```

```{r}
#Null states
nullmonths<-lapply(int.M,function(x){
  nullframe<-rand1(int,100)
  nullframe$Date_C<-unique(x$Date_C)
  return(nullframe)
})

#bind
nullmonths<-rbind_all(nullmonths)

nullmonths$Date_C<-as.POSIXct(nullmonths$Date_C,tz="")
nullmonths$variable<-NULL
colnames(nullmonths)<-c("Hummingbird","Mean_Null","Iteration","Date_C")
colnames(truemonths)<-c("Hummingbird","Mean_True","Lower_True","Higher_True","Date_C")

ggplot(nullmonths) + geom_violin(alpha=.8,aes(y=Mean_Null,x=Date_C,group=Date_C))+ geom_point(data=truemonths,aes(y=Mean_True,x=Date_C),col="red",size=1) + theme_bw() + ylab("Corolla Length Selected") + facet_wrap(~Hummingbird,scales="free") + scale_x_datetime(date_breaks = "6 months")

```

##Phase plot

Color code the randomization to tell when it is greater than expected, less than expected and even.

```{r}
#Standardize names
allmonths<-merge(nullmonths,truemonths,by=c("Hummingbird","Date_C"))

allmonths$Mean_Diff<-allmonths$Mean_Null-allmonths$Mean_True

#Month
allmonths$month<-months(as.POSIXct(allmonths$Date_C))
allmonths$month<-as.factor(allmonths$month)
allmonths$month<-factor(allmonths$month,levels=month.name)

#Year
allmonths$year<-years(as.POSIXct(allmonths$Date_C))

ggplot(allmonths) + geom_violin(alpha=.8,aes(y=Mean_Diff,x=month,col=as.factor(year),fill=Mean_Diff>0)) + theme_bw() + ylab("Difference between Bill and Corolla Length") + facet_wrap(~Hummingbird,scales="free",ncol=4)  + labs(fill="Evidence of Trait-matching") + xlab("Month")
```

As an assemblage
```{r}
ggplot(allmonths) + geom_violin(alpha=.8,aes(y=Mean_Diff,x=month,col=year,fill=Mean_Diff>0)) + theme_bw() + ylab("Difference between Bill and Corolla Length")   + labs(fill="Evidence of Trait-matching") 
```


As line plot

```{r}
sumnull<-allmonths %>% group_by(Hummingbird,month,year) %>% summarize(mean=mean(Mean_Diff),lower=quantile(Mean_Diff,0.05),upper=quantile(Mean_Diff,0.95))

ggplot(sumnull,aes(x=month)) + geom_line(alpha=.9,aes(y=mean,col=mean>0,group=year),size=1.2) + geom_linerange(aes(ymin=lower,ymax=upper)) + theme_bw() + ylab("Difference between Bill and Corolla Length") + facet_wrap(~Hummingbird,scales="free") + scale_color_manual("Evidence for Trait-matching",values=c("grey80","black")) 
```

##As a function of environment

```{r}
#Merge Flower Data
full.fl<-read.csv("InputData/FlowerTransectClean.csv")[,-1]

#month should be capital 
colnames(full.fl)[colnames(full.fl) %in% "month"]<-"Month"

#Take out flowers not in the network
full.fl<-full.fl[full.fl$Iplant_Double %in% levels(int$Iplant_Double),]

#Time Stamp
full.fl$Date_F<-as.POSIXct(full.fl$Date_F)

#Cut Stamp
full.fl$Date_C<-cut(full.fl$Date_F,breaks=unique(allmonths$Date_C))

#group by date cut
flower.month<-group_by(full.fl,Date_C,Transect_R,Date) %>% dplyr::summarise(Flowers=sum(Total_Flowers,na.rm=TRUE))%>% filter(Flowers>10) %>% summarize(Flowers=mean(Flowers)) %>% summarize(Flowers=sum(Flowers))

full.fl$Date_C<-as.POSIXct(full.fl$Date_C,tz = "")

#combine the flower totals and network metrics
flower.month$Date_C<-as.character(flower.month$Date_C)
allmonths$Date_C<-as.character(allmonths$Date_C)

network.fl<-merge(flower.month,allmonths,by="Date_C")
```

```{r}
ggplot(network.fl,aes(x=Flowers,y=Mean_Diff)) + geom_point() + facet_wrap(~Hummingbird,scales="free") + stat_smooth(method="lm")
```


Save image

```{r}

save.image("Randomization.Rdata")
```
