---
title: "Trait-matching and Available Resources: Randomizations"
author: "Ben Weinstein - Stony Brook University"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
    keep_md: true
  word_document: default
---

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(foreach)
library(doSNOW)
library(chron)
library(ggplot2)
library(knitr)
library(R2jags)
library(dplyr)
library(stringr)
library(gridExtra)
library(boot)
library(picante)
library(bipartite)
library(VennDiagram)

opts_chunk$set(message=FALSE,warning=FALSE,fig.width=9,fig.height=7,echo=TRUE,cache=F,fig.align='center',fig.path="figureObserved/")

set.seed(3)
#source functions
```

```{r,echo=F,cache=FALSE}
paste("Run Completed at",Sys.time())
```

```{r}
#reload if needed
load("Randomization.Rdata")
```

#Load in data

```{r}
int<-read.csv("InputData/FinalObs.csv")
```

#Randomization 1

##Is the observed distribution of choosen corollas different than the background distribution of availabile corollas.

This is most common definition of specialization.

```{r}
#true state
true_state<- indat %>% filter(State=="Used") %>% group_by(Hummingbird) %>% summarize(mean=mean(TotalCorolla))

mtruestate<-melt(true_state,measure.vars=c("mean"))

nullc<-function(indat){
  indat$State<-sample(indat$State)
null_state<- indat %>% filter(State=="Used") %>% group_by(Hummingbird) %>% summarize(mean=mean(TotalCorolla))

  return(null_state)
}

cl<-makeCluster(5,"SOCK")
registerDoSNOW(cl)
nullframe<-foreach(x=1:10000,.packages="dplyr") %dopar% {nullc(indat)}
stopCluster(cl)

names(nullframe)<-1:10000
nullframe<-melt(nullframe,id.vars=c("Hummingbird"))
```

#Compare null and observed distributions

## Mean

```{r}
ggplot(nullframe[nullframe$variable=="mean",],aes(x=value)) + geom_density(fill="black",alpha=.2) + geom_vline(data=mtruestate[mtruestate$variable=="mean",],aes(xintercept=value),linetype="dashed",col="red",size=1) + facet_wrap(~Hummingbird,scales="free") + theme_bw() 
```

Are those true distributions within reasonable confidence bands?

```{r}
sdat<-split(nullframe,nullframe$variable)
tstate<-sapply(sdat,function(d){
  sp<-split(d,d$Hummingbird,drop = T)
  l<-sapply(sp,function(y){
    #true state
    ts<-true_state[true_state$Hummingbird %in% unique(y$Hummingbird),as.character(unique(d$variable))]
    f<-ecdf(y$value)
    f(ts[[1]])
  })
})

#melt for plotting
mtstate<-melt(tstate)

ord<-hum.morph %>% arrange(desc(Bill)) %>% select(English) %>% .$English

mtstate$Var1<-factor(mtstate$Var1,levels=ord)

p<-ggplot(mtstate,aes(x=Var1,y=Var2,fill=(value<0.05 | value > 0.95))) + geom_tile() + coord_flip() + labs(x="Hummingbird",y="Metric",fill="Difference between\nObserved and Null Distribution\n(alpha=0.05)") + ggtitle("Used versus Background")
p
```

#Randomization II


```{r}
#true state
true_state<- indat %>% filter(State=="Used") %>% group_by(Hummingbird) %>% summarize(mean=mean(Traitmatch))

mtruestate<-melt(true_state,measure.vars=c("mean"))

#randomization function
nullc<-function(indat){
  indat$State<-sample(indat$State)
null_state<- indat %>% filter(State=="Used") %>% group_by(Hummingbird) %>% summarize(mean=mean(Traitmatch))

  return(null_state)
}

cl<-makeCluster(5,"SOCK")
registerDoSNOW(cl)
nullframe<-foreach(x=1:10000,.packages="dplyr") %dopar% {nullc(indat)}
stopCluster(cl)

names(nullframe)<-1:10000
nullframe<-melt(nullframe,id.vars=c("Hummingbird"))
```

#Compare null and observed distributions

## Mean Trait-matching

```{r}
ggplot(nullframe[nullframe$variable=="mean",],aes(x=value)) + geom_density(fill="black",alpha=.2) + geom_vline(data=mtruestate[mtruestate$variable=="mean",],aes(xintercept=value),linetype="dashed",col="red",size=1) + facet_wrap(~Hummingbird,scales="free") + theme_bw() 
```

Are those true distributions within reasonable confidence bands?

```{r}
sdat<-split(nullframe,nullframe$variable)
tstate<-sapply(sdat,function(d){
  sp<-split(d,d$Hummingbird,drop = T)
  l<-sapply(sp,function(y){
    #true state
    ts<-true_state[true_state$Hummingbird %in% unique(y$Hummingbird),as.character(unique(d$variable))]
    f<-ecdf(y$value)
    f(ts[[1]])
  })
})

#melt for plotting
mtstate<-melt(tstate)

ord<-hum.morph %>% arrange(desc(Bill)) %>% select(English) %>% .$English

mtstate$Var1<-factor(mtstate$Var1,levels=ord)

p2<-ggplot(mtstate[mtstate$Var2 == 'mean',],aes(x=Var1,y=Var2,fill=(value<0.05))) + geom_tile() + coord_flip() + labs(x="Hummingbird",y="Metric",fill="Difference between\nObserved and Null Distribution\n(alpha=0.05)") + ggtitle("Used versus Background")
p2
```

#Randomization III

Do the distributions of corollas used differ in resource periods, given the background in each time period?

Do they expand, contract, or shift?

```{r}
ggplot(indat[indat$State=="Used",],aes(x=TotalCorolla,fill=as.factor(Availability))) + geom_density(alpha=.7) + facet_wrap(~Hummingbird,scales="free",ncol=3) + theme_bw() + scale_fill_manual("Resource Availability",labels=c("Low","High"),values=c("Red","Black")) + ggtitle("Visited Flowers") + geom_vline(data=b,aes(xintercept=Bill),linetype="dashed",size=1,col='red') + scale_fill_manual("Resource Availability",values=c("grey90","black"),labels=c("Low","High")) + xlab("Flower Corolla Length (mm)") + xlim(0,65)
```

##True state of trait-matching among high and low periods.
```{r}
true_state<- indat %>% filter(State=="Used") %>% group_by(Hummingbird,Availability) %>% summarize(mean=mean(TotalCorolla),Lower=quantile(TotalCorolla,0.05),Higher=quantile(TotalCorolla,0.95)) 

mtruestate<-melt(true_state,measure.vars=c("mean","Higher","Lower"))

#difference in values
splt<-split(mtruestate,list(true_state$Hummingbird,mtruestate$variable),drop=T)

diffdf<-lapply(splt,function(x){
  cx<-acast(x,Availability~variable,value.var="value")
  return(data.frame(Hummingbird=unique(x$Hummingbird),Metric=unique(x$variable),Difference=diff(cx)[[1]]))
  })

truedf<-rbind_all(diffdf)
```

```{r}
#Define randomization function

nullc<-function(indat){
  
  #randomize 
  indat$Availability<- sample(indat$Availability)
  indat$State<-sample(indat$State)
  
  #Calculate null
true_state<- indat %>% filter(State=="Used") %>% group_by(Hummingbird,Availability) %>% summarize(mean=mean(TotalCorolla),Lower=quantile(TotalCorolla,0.05),Higher=quantile(TotalCorolla,0.95))

mtruestate<-melt(true_state,measure.vars=c("mean","Lower","Higher"))

#difference in values
splt<-split(mtruestate,list(true_state$Hummingbird,mtruestate$variable),drop=T)

diffdf<-lapply(splt,function(x){
  cx<-acast(x,Availability~variable,value.var="value")
  return(data.frame(Hummingbird=unique(x$Hummingbird),Metric=unique(x$variable),Difference=diff(cx)[[1]]))
  })

null_state<-rbind_all(diffdf)

  return(null_state)
}

cl<-makeCluster(5,"SOCK")
registerDoSNOW(cl)
nullframe<-foreach(x=1:10000,.packages=c("dplyr","reshape2")) %dopar% {nullc(indat)}
stopCluster(cl)

nullframe<-rbind_all(nullframe)
```

## Mean Difference in Corollas

```{r}
ggplot(data=nullframe[nullframe$Metric=="mean",],aes(x=Difference)) + geom_density(fill="black",alpha=.2) + geom_vline(data=truedf[truedf$Metric=="mean",],aes(xintercept=Difference),linetype="dashed",col="red")+ facet_wrap(~Hummingbird,scales="free") + theme_bw() + ggtitle("Is the mean corolla length different in high and low periods")
```


Are those true distributions within reasonable confidence bands?

```{r}
sdat<-split(nullframe,nullframe$Metric)
tstate<-sapply(sdat,function(d){
  sp<-split(d,d$Hummingbird,drop=T)
  l<-sapply(sp,function(y){
    #true state
    ts<-truedf[truedf$Hummingbird %in% unique(y$Hummingbird) & truedf$Metric %in% unique(y$Metric),"Difference"]
    
    f<-ecdf(y$Difference)
    f(ts[[1]])
  })
})

#melt for plotting
mtstate<-melt(tstate)

ord<-hum.morph %>% arrange(desc(Bill)) %>% select(English) %>% .$English

mtstate$Var1<-factor(mtstate$Var1,levels=ord)

#round to alpha interval
round(mtstate$value,2)
#order by billsize
p3<-ggplot(mtstate[,],aes(x=Var1,y=Var2,fill=(value<0.05 | value > 0.95))) + geom_tile() + coord_flip() + labs(x="Hummingbird",y="Metric",fill="Difference between\nObserved and Null Distribution\n(alpha=0.05)") + ggtitle("Used v Used")
p3
```

#View all randomization tests.
*
```{r}
grid.arrange(p,p2,p3,ncol=1)
```

Regardless of identity, does the community converge on a corolla length, what is the space used in each time period

```{r}
ggplot(indat[indat$State=="Used",],aes(fill=as.factor(Availability),x=TotalCorolla)) + geom_density(alpha=.75)+ scale_fill_manual("Resource Availability",values = c("grey90","black"),labels=c("Low","High")) + theme_bw() + geom_vline(xintercept=c(16.23,19.78),linetype="dashed") 
ggsave("Figures/AllCorollas.svg",height=4,width=6)
```

```{r}
#Order by bill length
indat$Hummingbird<-factor(indat$Hummingbird,levels=as.character(b[order(b$Bill),]$Hummingbird))

#loop through plots, by bill length.
b$Billc<-cut(b$Bill,breaks=c(0,16,30,60),labels=c("Short Bill","Medium Bill","Long Bill"))

indat<-merge(indat,b[,!colnames(b) %in% "Bill"])
```

```{r}
#loop through plots
p<-list()

#do we have enough senses.
m<-melt(table(indat$State,indat$Hummingbird,indat$Availability))

#take out fawn breasted brilliant
indat<-indat[!indat$Hummingbird %in% "Fawn-breasted Brilliant",]
#split data

sindat<-split(indat,indat$Billc)

#vectors of titles
for (x in 1:length(sindat)){
  j<-sindat[[x]]
  p[[x]]<-ggplot(j[j$State=="Used",],aes(x=TotalCorolla,fill=as.factor(Availability))) + geom_density(alpha=.7) + facet_wrap(~Hummingbird,scale="free_y",nrow=1) + theme_bw()   + geom_vline(aes(xintercept=Bill),linetype="dashed",size=1,col='red') + labs(x=NULL) + scale_fill_manual("Resource Availability",values=c("grey90","black"),labels=c("Low","High"),guide="none")+ theme(plot.margin=unit(c(.6,.6,0.1,0.1),"cm"))
}

#just add legend to last one.
p[[3]]<-p[[3]]+theme(legend.position='bottom') + scale_fill_manual("Resource Availability",values=c("grey90","black"),labels=c("Low","High"))
svg("Figures/ChangeCorolla_used.svg",height=8,width=10)
grid.arrange(p[[1]],p[[2]],p[[3]],heights=c(0.3,.3,0.4))
dev.off()

grid.arrange(p[[1]],p[[2]],p[[3]],heights=c(0.3,.3,0.4))

```

#Mean Shift

# Quantile shift

```{r}
# need a clever trick of where to play y axis to offset availability.
true_state$ymin<-0.05
true_state$ymax<-0.0515

true_state$ymin[true_state$Availability=="High"]<-true_state$ymin[true_state$Availability=="High"] + 0.005
true_state$ymax[true_state$Availability=="High"]<-true_state$ymax[true_state$Availability=="High"] + 0.005

#position of the means
tom<-true_state %>% group_by(Availability) %>% summarize(ymean=mean(c(ymin,ymax))+0.0001)

true_state<-merge(true_state,tom,by="Availability")
#plot distributions and ranges.

ggplot(indat[indat$State=='Unused',],aes(x=TotalCorolla,fill=Availability)) + geom_density(alpha=.7) + facet_wrap(~Hummingbird) + geom_rect(data=true_state[!true_state$Hummingbird %in% "Fawn-breasted Brilliant",],aes(ymin=ymin,ymax=ymax,xmin=Lower,xmax=Higher,x=NULL)) + scale_fill_manual("Resource Availability",values=c("blue","red"),labels=c("Low","High")) + scale_color_manual("Resource Availability",values=c("blue","red"),labels=c("Low","High"),guide="none") + theme_bw() + xlab("Corolla Length (mm)") + geom_vline(data=b[!b$Hummingbird %in% "Fawn-breasted Brilliant",],aes(xintercept=Bill),alpha=.3,size=1,linetype='dashed') + geom_point(data=true_state[!true_state$Hummingbird %in% "Fawn-breasted Brilliant",],size=3,aes(col=Availability,x=mean,y=ymean),col="black",show.legend=F) 

ggsave("Figures/Meanshift.jpeg",height=6,width=10,dpi=500)

```


Save image

```{r}

save.image("Randomization.Rdata")
```
